---
title: "Bubble_Plot"
output: html_document
date: "2025-12-15"
---

```{r setup, include=FALSE}

#Run only in chunk:
knitr::opts_knit$set(root.dir = '/Users/robertvandervelde/Library/CloudStorage/GoogleDrive-robertvandervelde@sydshafferlab.com/My Drive/Vander_Velde_Share/Experiments/RV143')

getwd()

```

#Load Libraries:
```{r, include=FALSE}

rm(list = ls())
library(ggplot2)
library(reshape2)


```


#Set variable:
```{r, include=FALSE}

font_size <- 8
font_type <- "Helvetica Neue"

```

#Load the gDNA barcode data:
```{r, include=FALSE}

gDNA_counts <- read.csv("2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Organized_Data/counts_without_spike_ins.csv", row.names = 1)

#Load spike-ins:
spike_in_counts <- read.csv("2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Organized_Data/spike_in_counts.csv", row.names = 1)

threshold_info <- read.csv("2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Organized_Data/threshold_type_dataframe.csv", row.names = 1)


```



#Get proportions of gDNA barcodes:
```{r, include=FALSE}

gDNA_proportion <- sweep(gDNA_counts, 2, colSums(gDNA_counts), "/")

```

#Get lineage barcodes that pass the 200 (>=) threshold in at least one group, using pools separately:
```{r, include=FALSE}

gDNA_proportion_200 <- gDNA_proportion[rowSums(threshold_info == "Two_Hundred") >= 1, ]

#Get pools separately:
gDNA_proportion_200_posttreatment_pool_1 <- gDNA_proportion[,c("Pre1","X1A","X1B")]
  
gDNA_proportion_200_posttreatment_pool_2 <- gDNA_proportion[,c("Pre2","X2A","X2B")]

#Get thresholded in just post-treatment samples:
gDNA_proportion_200_posttreatment_pool_1 <- gDNA_proportion_200_posttreatment_pool_1[rowSums(threshold_info[,c("X1A","X1B")] == "Two_Hundred") >= 1, ]

gDNA_proportion_200_posttreatment_pool_2 <- gDNA_proportion_200_posttreatment_pool_2[rowSums(threshold_info[,c("X2A","X2B")] == "Two_Hundred") >= 1, ]

#

#Look at overlapping barcodes from both datasets:
print(intersect(rownames(gDNA_proportion_200_posttreatment_pool_1), rownames(gDNA_proportion_200_posttreatment_pool_2)))

```


#Melt data frames and reorder based on pre1
```{r, include=FALSE}

gDNA_proportion_200_posttreatment_pool_1$Lineage <- rownames(gDNA_proportion_200_posttreatment_pool_1) 
gDNA_proportion_200_posttreatment_pool_2$Lineage <- rownames(gDNA_proportion_200_posttreatment_pool_2)

gDNA_proportion_200_posttreatment_pool_1_melt <- melt(gDNA_proportion_200_posttreatment_pool_1)
gDNA_proportion_200_posttreatment_pool_2_melt <- melt(gDNA_proportion_200_posttreatment_pool_2)

colnames(gDNA_proportion_200_posttreatment_pool_1_melt) <- c("Lineage","Source", "Proportion")
colnames(gDNA_proportion_200_posttreatment_pool_2_melt) <- c("Lineage","Source", "Proportion")

lineage_order <- gDNA_proportion_200_posttreatment_pool_1_melt %>%
  filter(Source == "Pre1") %>%
  arrange(Proportion) %>%
  pull(Lineage)

# Reorder the Lineage factor based on the calculated order
gDNA_proportion_200_posttreatment_pool_1_melt$Lineage <- 
  factor(gDNA_proportion_200_posttreatment_pool_1_melt$Lineage, 
         levels = lineage_order)

lineage_order <- gDNA_proportion_200_posttreatment_pool_2_melt %>%
  filter(Source == "Pre2") %>%
  arrange(Proportion) %>%
  pull(Lineage)

# Reorder the Lineage factor based on the calculated order
gDNA_proportion_200_posttreatment_pool_2_melt$Lineage <- 
  factor(gDNA_proportion_200_posttreatment_pool_2_melt$Lineage, 
         levels = lineage_order)




```



#Create a bubble plot pool 1:

```{r}

ggplot(gDNA_proportion_200_posttreatment_pool_1_melt, aes(x = Source, y = Lineage, size = Proportion)) + 
      geom_point(shape = 21, fill = NA) + 
      scale_size(range = c(.1, 24)) + 
      theme_minimal() +
      theme(
        legend.key = element_blank(),
        axis.text.x = element_text(vjust = 1),
        axis.text.y = element_blank(),
        panel.grid.major = element_line(color = "gray90"),
        panel.grid.minor = element_blank()
      ) + 
      labs(
        x = "", 
        y = "", 
        size = "Proportion of Reads", 
        fill = ""
      )

```


#Create a bubble plot, pool 2:

```{r}

ggplot(gDNA_proportion_200_posttreatment_pool_2_melt, aes(x = Source, y = Lineage, size = Proportion)) + 
      geom_point(shape = 21, fill = NA) + 
      scale_size(range = c(.1, 24)) + 
      theme_minimal() +
      theme(
        legend.key = element_blank(),
        axis.text.x = element_text(vjust = 1),
        axis.text.y = element_blank(),
        panel.grid.major = element_line(color = "gray90"),
        panel.grid.minor = element_blank()
      ) + 
      labs(
        x = "", 
        y = "", 
        size = "Proportion of Reads", 
        fill = ""
      )

```


#Get dataframe normalized by thresholded barcodes only:
```{r, include=FALSE}

#Get lineages 200 cells or more in pool 1 for counts:
gDNA_counts_200_pool_1 <- gDNA_counts[rownames(gDNA_proportion_200_posttreatment_pool_1),c("Pre1","X1A","X1B")]
#Normalize dataframe pool 1:
gDNA_counts_200_pool_1_normalized <- sweep(gDNA_counts_200_pool_1, 2, colSums(gDNA_counts_200_pool_1), "/")
#Melt dataframe, pool 1:
gDNA_counts_200_pool_1_normalized$Lineage <- rownames(gDNA_counts_200_pool_1)
gDNA_counts_200_pool_1_normalized_melted <- melt(gDNA_counts_200_pool_1_normalized)
colnames(gDNA_counts_200_pool_1_normalized_melted) <- c("Lineage","Source","Proportion")

#Put lineages in order of pre for pool 1:
lineage_order <- gDNA_counts_200_pool_1_normalized_melted %>%
  filter(Source == "Pre1") %>%
  arrange(Proportion) %>%
  pull(Lineage)

# Reorder the Lineage factor based on the calculated order
gDNA_counts_200_pool_1_normalized_melted$Lineage <- 
  factor(gDNA_counts_200_pool_1_normalized_melted$Lineage, 
         levels = lineage_order)



#Get lineages 200 cells or more in pool 2 for counts:
gDNA_counts_200_pool_2 <- gDNA_counts[rownames(gDNA_proportion_200_posttreatment_pool_2),c("Pre2","X2A","X2B")]
#Normalize dataframe pool 2:
gDNA_counts_200_pool_2_normalized <- sweep(gDNA_counts_200_pool_2, 2, colSums(gDNA_counts_200_pool_2), "/")
#Melt dataframe, pool 2:
gDNA_counts_200_pool_2_normalized$Lineage <- rownames(gDNA_counts_200_pool_2)
gDNA_counts_200_pool_2_normalized_melted <- melt(gDNA_counts_200_pool_2_normalized)
colnames(gDNA_counts_200_pool_2_normalized_melted) <- c("Lineage","Source","Proportion")

#Put lineages in order of pre for pool 1:
lineage_order <- gDNA_counts_200_pool_2_normalized_melted %>%
  filter(Source == "Pre2") %>%
  arrange(Proportion) %>%
  pull(Lineage)

# Reorder the Lineage factor based on the calculated order
gDNA_counts_200_pool_2_normalized_melted$Lineage <- 
  factor(gDNA_counts_200_pool_2_normalized_melted$Lineage, 
         levels = lineage_order)


```


#Create a bubble plot pool 1, normalized just to thresholded barcodes:

```{r}

rainbow_colors <- c("#FF0000", "#FF7F00", "#00FF00", "#0000FF", "#8B00FF")

#Number of lineages for rainbow:
n_lineages <- length(unique(gDNA_counts_200_pool_1_normalized_melted$Lineage))


ggplot(gDNA_counts_200_pool_1_normalized_melted, aes(x = Source, y = Lineage, size = Proportion)) + 
      geom_point(shape = 21, fill = NA, aes(color = as.numeric(factor(Lineage)))) + 
      scale_size(range = c(.1, 24)) + 
      scale_color_gradientn(colors = rep(rainbow_colors, length.out = n_lineages), guide = "none") +
      theme_minimal() +
      theme(
        legend.key = element_blank(),
        axis.text.x = element_text(vjust = 0.5, angle = 90, hjust = 1),
        axis.text.y = element_blank(),
        panel.grid.major = element_line(color = "gray90"),
        panel.grid.minor = element_blank(),
        text = element_text(family = font_type, size = font_size, color = "black"),
        axis.text = element_text(family = font_type, size = font_size, color = "black")
      ) +
      scale_x_discrete(labels = c("Pre1" = "Pre-Treatment  ", "X1A" = "Post-Treatment, Rep. A  ", "X1B" = "Post-Treatment, Rep. B  ")) +
      labs(
        x = "", 
        y = "",
        size = "Proportion of Reads", 
        fill = ""
      ) +
      #Include bubbles outside the panel/plotting area:
      coord_cartesian(clip = "off")


ggsave(filename = "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Plots/Bubble_Plot/Bubble_Plot_Pool1.pdf",
       device = cairo_pdf,
       width = 3, 
       height = 6)

```

#Create a bubble plot pool 2, normalized just to thresholded barcodes:

```{r}


rainbow_colors <- c("#FF0000", "#FF7F00", "#00FF00", "#0000FF", "#8B00FF")

#Number of lineages for rainbow:
n_lineages <- length(unique(gDNA_counts_200_pool_2_normalized_melted$Lineage))

ggplot(gDNA_counts_200_pool_2_normalized_melted, aes(x = Source, y = Lineage, size = Proportion)) + 
      geom_point(shape = 21, fill = NA, aes(color = as.numeric(factor(Lineage)))) + 
      scale_size(range = c(.1, 24)) + 
      scale_color_gradientn(colors = rep(rainbow_colors, length.out = n_lineages), guide = "none") +
      theme_minimal() +
      theme(
        legend.key = element_blank(),
        axis.text.x = element_text(vjust = 0.5, angle = 90, hjust = 1),
        axis.text.y = element_blank(),
        axis.ticks.length.x = unit(1, "in"),
        panel.grid.major = element_line(color = "gray90"),
        panel.grid.minor = element_blank(),
        text = element_text(family = font_type, size = font_size, color = "black"),
        axis.text = element_text(family = font_type, size = font_size, color = "black")
      ) +
      scale_x_discrete(labels = c("Pre2" = "Pre-Treatment  ", "X2A" = "Post-Treatment, Rep. A  ", "X2B" = "Post-Treatment, Rep. B  ")) +
      labs(
        x = "", 
        y = "",
        size = "Proportion of Reads", 
        fill = ""
      ) +
      #Include bubbles outside the panel/plotting area:
      coord_cartesian(clip = "off")


ggsave(filename = "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Plots/Bubble_Plot/Bubble_Plot_Pool2.pdf",
       device = cairo_pdf,
       width = 3, 
       height = 6)

```
