---
title: "Diversity_Metrics"
output: html_document
date: "2025-10-29"
---


```{r setup, include=FALSE}

#Run only in chunk:
knitr::opts_knit$set(root.dir = '/Users/robertvandervelde/Library/CloudStorage/GoogleDrive-robertvandervelde@sydshafferlab.com/My Drive/Vander_Velde_Share/Experiments/RV143')

getwd()

```

#Load Libraries:
```{r, include=FALSE}

rm(list = ls())
library(tidyverse)
library(Matrix)
library(dplyr)
library(reshape2)
library(vegan)

library(egg)

```

#Set variable:
```{r, include=FALSE}

font_size <- 8
font_type <- "Helvetica Neue"

panel_dimensions <- c(0.45,1.5)
bar_width <- 0.5

read_threshold <- 2

```

#Define functions:
```{r, include=FALSE}

#Define revcomp function:
revcomp <- function(seq) {
  chartr("ATCG", "TAGC", paste(rev(strsplit(seq, "")[[1]]), collapse=""))
}

```


#Load the gDNA barcode data:
```{r, include=FALSE}

gDNA_counts <- read.csv("2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Organized_Data/counts_without_spike_ins.csv", row.names = 1)

#Get reverse complement for barcode names:
rownames(gDNA_counts) <- sapply(rownames(gDNA_counts), revcomp)

#Load spike-ins:
spike_in_counts <- read.csv("2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Organized_Data/spike_in_counts.csv", row.names = 1)

#Get reverse complement for barcode names:
rownames(spike_in_counts) <- sapply(rownames(spike_in_counts), revcomp)


```

#Get thresholded data for each pool separately:
```{r, include=FALSE}

gDNA_counts_thresholded_pool1 <- gDNA_counts[rowSums(gDNA_counts[,c("Pre1","X1A","X1B")] >= read_threshold)>0, c("Pre1","X1A","X1B")]

gDNA_counts_thresholded_pool2 <- gDNA_counts[rowSums(gDNA_counts[,c("Pre2","X2A","X2B")] >= read_threshold)>0, c("Pre2","X2A","X2B")]


gDNA_proportion_thresholded_pool1 <- sweep(gDNA_counts_thresholded_pool1, 2, colSums(gDNA_counts_thresholded_pool1), "/")

gDNA_proportion_thresholded_pool2 <- sweep(gDNA_counts_thresholded_pool2, 2, colSums(gDNA_counts_thresholded_pool2), "/")


```


#Calculate diversity metrics separately for pool 1 and pool 2 using thresholded data:
```{r, include=FALSE}

#Calculate number of lineage barcodes:
number_of_unique_barcodes_thresholded_pool1 <- colSums(gDNA_counts_thresholded_pool1 != 0)
number_of_unique_barcodes_thresholded_pool1 <- melt(number_of_unique_barcodes_thresholded_pool1)
number_of_unique_barcodes_thresholded_pool1$Source <- rownames(number_of_unique_barcodes_thresholded_pool1)

number_of_unique_barcodes_thresholded_pool2 <- colSums(gDNA_counts_thresholded_pool2 != 0)
number_of_unique_barcodes_thresholded_pool2 <- melt(number_of_unique_barcodes_thresholded_pool2)
number_of_unique_barcodes_thresholded_pool2$Source <- rownames(number_of_unique_barcodes_thresholded_pool2)

#Calculate Shannon diversity:
shannon_diversity_thresholded_pool1 <- sapply(gDNA_proportion_thresholded_pool1, function(x) vegan::diversity(x, index = "shannon"))
shannon_diversity_thresholded_pool1 <- melt(shannon_diversity_thresholded_pool1)
shannon_diversity_thresholded_pool1$Source <- rownames(shannon_diversity_thresholded_pool1)

shannon_diversity_thresholded_pool2 <- sapply(gDNA_proportion_thresholded_pool2, function(x) vegan::diversity(x, index = "shannon"))
shannon_diversity_thresholded_pool2 <- melt(shannon_diversity_thresholded_pool2)
shannon_diversity_thresholded_pool2$Source <- rownames(shannon_diversity_thresholded_pool2)

#Calculate Simpson diversity:
simpson_diversity_thresholded_pool1 <- sapply(gDNA_proportion_thresholded_pool1, function(x) vegan::diversity(x, index = "simpson"))
simpson_diversity_thresholded_pool1 <- melt(simpson_diversity_thresholded_pool1)
simpson_diversity_thresholded_pool1$Source <- rownames(simpson_diversity_thresholded_pool1)

simpson_diversity_thresholded_pool2 <- sapply(gDNA_proportion_thresholded_pool2, function(x) vegan::diversity(x, index = "simpson"))
simpson_diversity_thresholded_pool2 <- melt(simpson_diversity_thresholded_pool2)
simpson_diversity_thresholded_pool2$Source <- rownames(simpson_diversity_thresholded_pool2)

```


#Plot thesholded number of unique barcodes, pool 1:
```{r}

p <- ggplot(data = number_of_unique_barcodes_thresholded_pool1) +
  geom_bar(aes(x = Source, y = value, fill = Source), stat = "identity", width = bar_width) +
  scale_x_discrete(
    limits = c("Pre1","X1A","X1B"),
    labels = c("Pre1" = "Pre-Treatment", "X1A" = "Post-Treatment, Rep. A", "X1B" = "Post-Treatment, Rep. B")
  ) +
  scale_y_continuous(
    name = "# of Unique Lineage Barcodes",
    expand = expansion(mult = c(0,0.05))
  ) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_text(family = font_type, size = font_size, color = "black", angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", margin = margin(t = 0, r = 0.05, b = 0, l = 0)),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.position = "none"
  ) +
  scale_fill_manual(values = c("Pre1" = "red", "X1A" = "pink", "X1B" = "hotpink")) +
  xlab("")

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Plots/Diversity_Metrics/Number_of_Barcodes_Thresholded_Pool1.pdf", device = cairo_pdf, width = 3, height = 5)

```


#Plot thesholded number of unique barcodes, pool 2:
```{r}

p <- ggplot(data = number_of_unique_barcodes_thresholded_pool2) +
  geom_bar(aes(x = Source, y = value, fill = Source), stat = "identity", width = bar_width) +
  scale_x_discrete(
    limits = c("Pre2","X2A","X2B"),
    labels = c("Pre2" = "Pre-Treatment", "X2A" = "Post-Treatment, Rep. A", "X2B" = "Post-Treatment, Rep. B")
  ) +
  scale_y_continuous(
    name = "# of Unique Lineage Barcodes",
    expand = expansion(mult = c(0,0.05))
  ) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_text(family = font_type, size = font_size, color = "black", angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", margin = margin(t = 0, r = 0.05, b = 0, l = 0)),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.position = "none"
  ) +
  scale_fill_manual(values =  c("Pre2" = "blue", "X2A" ="turquoise", "X2B" = "cyan")) +
  xlab("")

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Plots/Diversity_Metrics/Number_of_Barcodes_Thresholded_Pool2.pdf", device = cairo_pdf, width = 3, height = 5)

```


#Plot thesholded Shannon diversity, pool 1:
```{r}

p <- ggplot(data = shannon_diversity_thresholded_pool1) +
  geom_bar(aes(x = Source, y = value, fill = Source), stat = "identity", width = bar_width) +
  scale_x_discrete(
    limits = c("Pre1","X1A","X1B"),
    labels = c("Pre1" = "Pre-Treatment", "X1A" = "Post-Treatment, Rep. A", "X1B" = "Post-Treatment, Rep. B")
  ) +
  scale_y_continuous(
    name = "Shannon Diversity Index",
    expand = expansion(mult = c(0,0.05))
  ) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_text(family = font_type, size = font_size, color = "black", angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", margin = margin(t = 0, r = 0.05, b = 0, l = 0)),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.position = "none"
  ) +
  scale_fill_manual(values = c("Pre1" = "red", "X1A" = "pink", "X1B" = "hotpink")) +
  xlab("")

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Plots/Diversity_Metrics/Shannon_Thresholded_Pool1.pdf", device = cairo_pdf, width = 3, height = 5)

```

#Plot thesholded Shannon diversity, pool 2:
```{r}

p <- ggplot(data = shannon_diversity_thresholded_pool2) +
  geom_bar(aes(x = Source, y = value, fill = Source), stat = "identity", width = bar_width) +
  scale_x_discrete(
    limits = c("Pre2","X2A","X2B"),
    labels = c("Pre2" = "Pre-Treatment", "X2A" = "Post-Treatment, Rep. A", "X2B" = "Post-Treatment, Rep. B")
  ) +
  scale_y_continuous(
    name = "Shannon Diversity Index",
    expand = expansion(mult = c(0,0.05))
  ) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_text(family = font_type, size = font_size, color = "black", angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", margin = margin(t = 0, r = 0.05, b = 0, l = 0)),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.position = "none"
  ) +
  scale_fill_manual(values =  c("Pre2" = "blue", "X2A" ="turquoise", "X2B" = "cyan")) +
  xlab("")

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Plots/Diversity_Metrics/Shannon_Thresholded_Pool2.pdf", device = cairo_pdf, width = 3, height = 5)

```


#Plot thesholded Simpson diversity, pool 1:
```{r}

p <- ggplot(data = simpson_diversity_thresholded_pool1) +
  geom_bar(aes(x = Source, y = value, fill = Source), stat = "identity", width = bar_width) +
  scale_x_discrete(
    limits = c("Pre1","X1A","X1B"),
    labels = c("Pre1" = "Pre-Treatment", "X1A" = "Post-Treatment, Rep. A", "X1B" = "Post-Treatment, Rep. B")
  ) +
  scale_y_continuous(
    name = "Simpson Diversity Index",
    expand = expansion(mult = c(0,0.05))
  ) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_text(family = font_type, size = font_size, color = "black", angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", margin = margin(t = 0, r = 0.05, b = 0, l = 0)),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.position = "none"
  ) +
  scale_fill_manual(values = c("Pre1" = "red", "X1A" = "pink", "X1B" = "hotpink")) +
  xlab("")

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Plots/Diversity_Metrics/Simpson_Thresholded_Pool1.pdf", device = cairo_pdf, width = 3, height = 5)

```


#Plot thesholded Simpson diversity, pool 2:
```{r}

p <- ggplot(data = simpson_diversity_thresholded_pool2) +
  geom_bar(aes(x = Source, y = value, fill = Source), stat = "identity", width = bar_width) +
  scale_x_discrete(
    limits = c("Pre2","X2A","X2B"),
    labels = c("Pre2" = "Pre-Treatment", "X2A" = "Post-Treatment, Rep. A", "X2B" = "Post-Treatment, Rep. B")
  ) +
  scale_y_continuous(
    name = "Simpson Diversity Index",
    expand = expansion(mult = c(0,0.05))
  ) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_text(family = font_type, size = font_size, color = "black", angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", margin = margin(t = 0, r = 0.05, b = 0, l = 0)),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.position = "none"
  ) +
  scale_fill_manual(values =  c("Pre2" = "blue", "X2A" ="turquoise", "X2B" = "cyan")) +
  xlab("")

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Plots/Diversity_Metrics/Simpson_Thresholded_Pool2.pdf", device = cairo_pdf, width = 3, height = 5)

```
