---
title: "Organize_Data"
output: html_document
date: "2024-12-04"
---

```{r setup, include=FALSE}

#Run only in chunk:
knitr::opts_knit$set(root.dir = '/Users/robertvandervelde/Library/CloudStorage/GoogleDrive-robertvandervelde@sydshafferlab.com/My Drive/Vander_Velde_Share/Experiments/RV143')

getwd()

```

#Load Libraries and delete variables:
```{r, include=FALSE}

getwd()

rm(list = ls())

library(ggplot2)
library(extrafont)
#font_import()
loadfonts()
library(ggrastr)

library(stringi)
library(stringr)

```

#Define functions:

```{r, include=FALSE}

#For reverse complementing:
revcomp <- function(seq) {
  chartr("ATCG", "TAGC", paste(rev(strsplit(seq, "")[[1]]), collapse=""))
}

```

#Load data:
```{r, include=FALSE}

data <- read.csv("2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/barcode_quantify/FeatureReference_filtered_group1_counts.csv", row.names = 6)

#Remove non-count rows from data:
data <- data[,-c(1:6)]

#Rename columns:
colnames(data) <- c("1A","1B","2A","2B","Pre1","Pre2")

#Reorder the columns:
data <- data[,c("Pre1","1A","1B","Pre2","2A","2B")]

```

#Load the spike-ins:
```{r, include=FALSE}

spike_ins <- read.csv("spike_in.csv")

```

#Add spike-ins to a new dataframe and create a dataframe without spike-ins:

```{r, include=FALSE}

#Get spike-in counts from dataframe:
df_spike_in_counts <- data[(rownames(data) %in% spike_ins$sequence), ]

#Rearrange so same order as in spike-in dataframe:
df_spike_in_counts <- data[c(spike_ins$sequence),]

plot(df_spike_in_counts$Pre1, df_spike_in_counts$Pre2, col=factor(df_spike_in_counts$Cell_Numbers), pch=16)
plot(df_spike_in_counts$"1A", df_spike_in_counts$"2A", col=factor(df_spike_in_counts$Cell_Numbers), pch=16)
plot(df_spike_in_counts$"1B", df_spike_in_counts$"2B", col=factor(df_spike_in_counts$Cell_Numbers), pch=16)

#Add cell numbers to spike-in cound dataframe:
df_spike_in_counts$Cell_Numbers <- spike_ins$spike_med

#Create dataframe without spike-ins:
df_remove_spike_ins <- data[!(rownames(data) %in% spike_ins$sequence), ]

```

#Save the dataframes for spike-in counts and non-spike-in counts:
```{r, include=FALSE}

write.csv(df_remove_spike_ins, "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Organized_Data/counts_without_spike_ins.csv")

write.csv(df_spike_in_counts, "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Organized_Data/spike_in_counts.csv")

```

#Get the means for all spike-ins numbers:
```{r, include=FALSE}

spike_in_20_means <- colMeans(df_spike_in_counts[df_spike_in_counts$Cell_Numbers==20,][c(1:6)])
spike_in_100_means <- colMeans(df_spike_in_counts[df_spike_in_counts$Cell_Numbers==100,][c(1:6)])
spike_in_200_means <- colMeans(df_spike_in_counts[df_spike_in_counts$Cell_Numbers==200,][c(1:6)])

```

#Divide barcodes into sections based on where they fall based on barcodes:
```{r, include=FALSE}

#Create dataframe with the same size as data without spike-ins:
spike_in_threshold_means <- data.frame(matrix(ncol = ncol(df_remove_spike_ins), nrow = nrow(df_remove_spike_ins)))
colnames(spike_in_threshold_means) <- colnames(df_remove_spike_ins)
rownames(spike_in_threshold_means) <- rownames(df_remove_spike_ins)

#Add strings to represent what they are greater than or equal too:
spike_in_threshold_means[] <- mapply(function(col, t20, t100, t200) 
  #Order is important to allow higher thresholds:
  ifelse(col >= t200, "Two_Hundred",
  ifelse(col >= t100, "One_Hundred",
  ifelse(col >= t20, "Twenty", 
  "None"))),
  df_remove_spike_ins, 
  spike_in_20_means, 
  spike_in_100_means, 
  spike_in_200_means)

df_remove_spike_ins[spike_in_threshold_means$"2A"=="Two_Hundred" & spike_in_threshold_means$"2B"=="Two_Hundred",]

```

#Save dataframe with threshold values:
```{r, include=FALSE}

write.csv(spike_in_threshold_means, "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Organized_Data/threshold_type_dataframe.csv")

```

#Make a simple plot showing threshold values for barcode pool 2:
```{r}

ggplot(df_remove_spike_ins, aes(x = `2A`, y = `2B`)) +
   geom_point_rast(raster.dpi = 300) +
   geom_vline(xintercept = spike_in_20_means["2B"], linetype = "dotted", color = "red") +
   geom_vline(xintercept = spike_in_100_means["2B"], linetype = "dotted", color = "blue") +
   geom_vline(xintercept = spike_in_200_means["2B"], linetype = "dotted", color = "orange") +
   geom_hline(yintercept = spike_in_20_means["2A"], linetype = "dotted", color = "red") +
   geom_hline(yintercept = spike_in_100_means["2A"], linetype = "dotted", color = "blue") +
   geom_hline(yintercept = spike_in_200_means["2A"], linetype = "dotted", color = "orange") +
   theme(text = element_text(family = "Helvetica Neue")) +
   coord_fixed(ratio = 1)

ggsave("2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Plots/Random/Barcode_Pool_2.pdf", plot = last_plot(), device = cairo_pdf, width = 8, height = 8)



```


#Create lists of interesting barcodes:
```{r, include=FALSE}

two_hundred_or_higher <- df_remove_spike_ins[spike_in_threshold_means$"2A"=="Two_Hundred" & spike_in_threshold_means$"2B"=="Two_Hundred",]

nrow(df_remove_spike_ins[spike_in_threshold_means$"2A"=="Two_Hundred",])-21

nrow(df_remove_spike_ins[spike_in_threshold_means$"2B"=="Two_Hundred",])-21

nrow(df_remove_spike_ins[spike_in_threshold_means$"1A"=="Two_Hundred",])

nrow(df_remove_spike_ins[spike_in_threshold_means$"2A"=="Two_Hundred",])

nrow(df_remove_spike_ins[spike_in_threshold_means$"1A"=="Two_Hundred" & spike_in_threshold_means$"2A"=="Two_Hundred",])


```

#Check if lists of interesting barcodes have correct pattern:
```{r, include=FALSE}

#Create pattern:
#Define weak, strong and any:
W <- c("A","T")
S <- c("C","G")
N <- c(W,S)

#Create the pattern, this is the reverse complement of the plasmid since we used the scaffold and older code for analysis:
pattern <- paste0("^",
                    "([", paste(W, collapse = ""), "]",
                    "[", paste(N, collapse = ""), "]",
                    "[", paste(S, collapse = ""), "]",
                    "[", paste(N, collapse = ""), "])",
                    "{5}$")

#Get list of booleans for matching and unmatching barcodes:
two_hundred_or_higher_with_pattern_boolean <- str_detect(row.names(two_hundred_or_higher), pattern)

#Get interesting barcodes that don't match the pattern:
two_hundred_or_higher_with_pattern_boolean_mismatch <- two_hundred_or_higher[!two_hundred_or_higher_with_pattern_boolean,]

```

#Create strings for making recovery barcodes (in plasmid VB230528-1280dkp.), this sequence is arguably too long as the annealing temperature >>50C, however, the effects of this are not clear. These barcodes do not all match the pattern:
```{r, include=FALSE}

#Get the reverse complement of barcodes, since this time we took from the scaffold:
revcomp_barcodes <-  sapply(row.names(two_hundred_or_higher), revcomp)

#Add the ends of the primers, increasing the overlap since oligos <121 are the same cost at TWIST:
gibson_assembly_sequences <- paste("TTATCGGATCAACTTTGTATAGAAAAGTTGTACTGTGGTATCACTCGTC", revcomp_barcodes, "AGGCTCAGCTAAGGTGCAATGGTAGGCGTGTACGGTGGGAGGCCTATATAA", sep = "")

#Save sequences for TWIST:
write.table(as.data.frame(gibson_assembly_sequences), 
          "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Organized_Data/Gibson_Assembly_Sequences_20241212.csv",
          row.names = FALSE, 
          col.names = FALSE,
          quote = FALSE,
          sep = ",")

```


#Create strings for making recovery barcodes for apoptosis plasmid (VB240701-2532bdu):
#These barcodes do not all match the pattern:
```{r, include=FALSE}

#Add the ends of the primers, aiming for just above 50C annealing temperature, ignoring overhang which is presumably digested first:
gibson_assembly_sequences_apoptosis <- paste("ATCAACTTTGTATAGAAAAGTTGTACTGTGGTATCACTCGTC", revcomp_barcodes, "AGGCTCAGCTAAGGTGCAATGGTAGGCGTGTACGGTG", sep = "")

#Save sequences for TWIST:
write.table(as.data.frame(gibson_assembly_sequences_apoptosis), 
          "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Organized_Data/Gibson_Assembly_Sequences_20250401_Apoptosis.csv",
          row.names = FALSE, 
          col.names = FALSE,
          quote = FALSE,
          sep = ",")

```

#Find barcodes with a similar number of reads as the "primed" cells above in Pre2:
```{r, include=FALSE}

#Get a sorted dataframe to easily find similar barcodes:
df_remove_spike_ins_Pre2_sorted <- df_remove_spike_ins[order(df_remove_spike_ins$Pre2, decreasing = TRUE), ]

#Get the indices of primed cells in the above dataframe:
primed_indices <- which(rownames(df_remove_spike_ins_Pre2_sorted) %in% rownames(two_hundred_or_higher))

#Get barcodes that are just above primed cells in Pre2:
barcodes_above_primed <- df_remove_spike_ins_Pre2_sorted[primed_indices-1,]

#Get barcodes that are just below primed cells in Pre2:
barcodes_below_primed <- df_remove_spike_ins_Pre2_sorted[primed_indices+1,]

#Combine every other barcode from above:
combined_above_below_primed <- rbind(barcodes_below_primed[rep(c(TRUE, FALSE), length.out = nrow(barcodes_below_primed)), ], barcodes_above_primed[!rep(c(TRUE, FALSE), length.out = nrow(barcodes_above_primed)), ])

#View percentage of reads in primed vs near primed:
sum(combined_above_below_primed$Pre2)/sum(df_remove_spike_ins$Pre2)
sum(two_hundred_or_higher$Pre2)/sum(df_remove_spike_ins$Pre2)

#Check if these barcodes match the NSNW pattern:
combined_above_below_primed_string_detect <- str_detect(row.names(combined_above_below_primed), pattern)

combined_above_below_primed_boolean_mismatch <- combined_above_below_primed[!combined_above_below_primed_string_detect,]

#In the primed screen I allowed barcodes that don't match the pattern, however here the barcode appears to be shifted, leading , I am therefore replacing it:

#Find index in barcodes_above_primed

mismatch_above_primed_index <- which(rownames(barcodes_above_primed) == rownames(combined_above_below_primed_boolean_mismatch)[1])

#Create new list replacing mismatch from barcode in barcodes_below_primed:
combined_above_below_primed_without_mismatch <- combined_above_below_primed

combined_above_below_primed_without_mismatch[rownames(combined_above_below_primed_boolean_mismatch)[1],] <- barcodes_below_primed[mismatch_above_primed_index,]

#Also change the replaced barcode:
combined_above_below_primed_without_mismatch[rownames(combined_above_below_primed_boolean_mismatch)[1],] <- barcodes_below_primed[mismatch_above_primed_index,]; rownames(combined_above_below_primed_without_mismatch)[which(rownames(combined_above_below_primed_without_mismatch) == rownames(combined_above_below_primed_boolean_mismatch)[1])] <- rownames(barcodes_below_primed)[mismatch_above_primed_index]

#Get reverse complement of above barcodes:
combined_above_below_primed_without_mismatch_revcomp <- sapply(row.names(combined_above_below_primed_without_mismatch), revcomp)


```

```{r, include=FALSE}

#Add the ends of the primers, aiming for just above 50C annealing temperature, ignoring overhang which is presumably digested first:
gibson_assembly_sequences_apoptosis_nonprimed <- paste("ATCAACTTTGTATAGAAAAGTTGTACTGTGGTATCACTCGTC", combined_above_below_primed_without_mismatch_revcomp, "AGGCTCAGCTAAGGTGCAATGGTAGGCGTGTACGGTG", sep = "")

#Save sequences for TWIST:
write.table(as.data.frame(gibson_assembly_sequences_apoptosis_nonprimed), 
          "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Organized_Data/Gibson_Assembly_Sequences_20250401_Apoptosis_Nonprimed.csv",
          row.names = FALSE, 
          col.names = FALSE,
          quote = FALSE,
          sep = ",")

```

#Get all barcodes that match the pattern:
```{r, include=FALSE}

#Get name of barcodes with pattern:
names_remove_spike_ins_match_pattern <- str_detect(row.names(df_remove_spike_ins), pattern)

#Create dataframe with only matching barcodes:
df_remove_spike_ins_match_pattern <- df_remove_spike_ins[names_remove_spike_ins_match_pattern,]

```


