---
title: "qd_diversity"
output: html_document
date: "2025-09-18"
---

```{r setup, include=FALSE}

#Run only in chunk:
knitr::opts_knit$set(root.dir = '/Users/robertvandervelde/Library/CloudStorage/GoogleDrive-robertvandervelde@sydshafferlab.com/My Drive/Vander_Velde_Share/Experiments/RV143')

getwd()

```

#Load Libraries:
```{r, include=FALSE}

rm(list = ls())
library(tidyverse)
library(Matrix)
library(dplyr)
library(reshape2)
library(vegan)

library(egg)

```

#Set variable:
```{r, include=FALSE}

font_size <- 8
font_type <- "Helvetica Neue"

panel_dimensions <- c(0.65,1.5)

read_threshold <- 2

```

#Define functions:
```{r, include=FALSE}

#Define revcomp function:
revcomp <- function(seq) {
  chartr("ATCG", "TAGC", paste(rev(strsplit(seq, "")[[1]]), collapse=""))
}

```


#Load the gDNA barcode data:
```{r, include=FALSE}

gDNA_counts <- read.csv("2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Organized_Data/counts_without_spike_ins.csv", row.names = 1)

#Get reverse complement for barcode names:
rownames(gDNA_counts) <- sapply(rownames(gDNA_counts), revcomp)

#Load spike-ins:
spike_in_counts <- read.csv("2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Organized_Data/spike_in_counts.csv", row.names = 1)

#Get reverse complement for barcode names:
rownames(spike_in_counts) <- sapply(rownames(spike_in_counts), revcomp)


```



#Calculate proportions of gDNA reads:
```{r, include=FALSE}

gDNA_proportion <- sweep(gDNA_counts, 2, colSums(gDNA_counts), "/")

```

#Calculate q_D for each sample:
```{r, include=FALSE}

#Set range of q values:
q_values <- c(10^seq(-2, 2, by = 0.001))

q_D_values <- sapply(names(gDNA_proportion), function(col_name) {
  sapply(q_values, function(q) {
    (sum(gDNA_proportion[[col_name]]^q))^(1/(1-q))
  })
})

#Create a dataframe with q_D values:
q_D_values_df <- as.data.frame(cbind(q_values, q_D_values))

#Melt dataframe for plotting (except for q)
q_D_values_df_melt <- melt(q_D_values_df, 
                           id.vars = "q_values",  
                           measure.vars = 2:dim(q_D_values_df)[2],  
                           variable.name = "Sample",
                           value.name = "qD")

colnames(q_D_values_df_melt) <- c("q", "Sample", "q_D")

#Remove q=1 values for log-log plotting:
q_D_values_df_melt_nonone_q <- q_D_values_df_melt[q_D_values_df_melt$q != 1, ]

```

#Creat plot for all q_D:
```{r}

ggplot(q_D_values_df_melt_nonone_q, aes(x = q, y = q_D, colour = Sample)) +
  geom_line(linewidth = 0.5) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.title = element_blank(),
    legend.key.height = unit(0.75, "lines")
  ) +
  xlab("q") +
  ylab("qD") +
  scale_x_log10() +
  scale_y_log10() +
  scale_color_manual(labels = c("Pre1" = "Pre-Treatment, Pool 1", "X1A" = "Post-Treatment, Pool 1, Replicate A", "X1B" = "Post-Treatment, Pool 1, Replicate B", "X2B" = "Post-Treatment, Pool 1, Replicate B", "Pre2" = "Pre-Treatment, Pool 2", "X2A" = "Post-Treatment, Pool 2, Replicate A", "X2B" = "Post-Treatment, Pool 2, Replicate B"), values = c("Pre1" = "red", "Pre2" = "blue", "X1A" = "pink", "X1B" = "hotpink", "X2A" ="turquoise", "X2B" = "cyan"))


ggsave("2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Plots/qd_diversity/q_D_lineplot.pdf", plot = last_plot(), device = cairo_pdf, width = 5, height = 2)


```
#Get exact barcode counts to check:
```{r, include=FALSE}

# Count non-zero elements for each column (excluding the first column)
non_zero_counts <- sapply(gDNA_counts, function(x) sum(x != 0, na.rm = TRUE))

# Create the dataframe with sample names and non-zero counts
df_number_of_barcodes <- data.frame(
  Sample = colnames(gDNA_counts),
  Non_Zero_Count = non_zero_counts
)

rm(non_zero_counts)

```

#Get individual diversity metrics, without thresholds:
```{r, include=FALSE}

#Exponent of shannon diversity:
exp(sapply(gDNA_proportion, function(x) vegan::diversity(x, index = "shannon")))

number_of_unique_barcodes <- colSums(gDNA_proportion != 0)
number_of_unique_barcodes <- melt(number_of_unique_barcodes)
number_of_unique_barcodes$Source <- rownames(number_of_unique_barcodes)

shannon_diversity <- sapply(gDNA_proportion, function(x) vegan::diversity(x, index = "shannon"))
shannon_diversity <- melt(shannon_diversity)
shannon_diversity$Source <- rownames(shannon_diversity)

simpson_diversity <- sapply(gDNA_proportion, function(x) vegan::diversity(x, index = "simpson"))
simpson_diversity <- melt(simpson_diversity)
simpson_diversity$Source <- rownames(simpson_diversity)

```

#Plot untrhesholded number of unique barcodes:
```{r}

p <- ggplot(data = number_of_unique_barcodes) +
  geom_bar(aes(x = Source, y = value, fill = Source), stat = "identity", width = 0.5) +
  scale_x_discrete(
    limits = c("Pre1","X1A","X1B","Pre2","X2A","X2B"),
    labels = c("Pre1" = "Pre-Treatment, Pool 1", "X1A" = "Post-Treatment, Pool 1, Replicate A", "X1B" = "Post-Treatment, Pool 1, Replicate B", "X2B" = "Post-Treatment, Pool 1, Replicate B", "Pre2" = "Pre-Treatment, Pool 2", "X2A" = "Post-Treatment, Pool 2, Replicate A", "X2B" = "Post-Treatment, Pool 2, Replicate B")
  ) +
  scale_y_continuous(
    name = "# of Unique Lineage Barcodes",
    expand = expansion(mult = c(0,0.05))
  ) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_text(family = font_type, size = font_size, color = "black", angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", margin = margin(t = 0, r = 0.05, b = 0, l = 0)),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.position = "none"
  ) +
  scale_fill_manual(values = c("Pre1" = "red", "Pre2" = "blue", "X1A" = "pink", "X1B" = "hotpink", "X2A" ="turquoise", "X2B" = "cyan")) +
  xlab("")

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Plots/qd_diversity/Number_of_Barcodes_Unthresholded.pdf", device = cairo_pdf, width = 3, height = 5)

```


#Plot untrhesholded Shannon diversity:
```{r}

p <- ggplot(data = shannon_diversity) +
  geom_bar(aes(x = Source, y = value, fill = Source), stat = "identity", width = 0.5) +
  scale_x_discrete(
    limits = c("Pre1","X1A","X1B","Pre2","X2A","X2B"),
    labels = c("Pre1" = "Pre-Treatment, Pool 1", "X1A" = "Post-Treatment, Pool 1, Replicate A", "X1B" = "Post-Treatment, Pool 1, Replicate B", "X2B" = "Post-Treatment, Pool 1, Replicate B", "Pre2" = "Pre-Treatment, Pool 2", "X2A" = "Post-Treatment, Pool 2, Replicate A", "X2B" = "Post-Treatment, Pool 2, Replicate B")
  ) +
  scale_y_continuous(
    name = "Shannon Diversity Index",
    expand = expansion(mult = c(0,0.05))
  ) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_text(family = font_type, size = font_size, color = "black", angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", margin = margin(t = 0, r = 0.05, b = 0, l = 0)),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.position = "none"
  ) +
  scale_fill_manual(values = c("Pre1" = "red", "Pre2" = "blue", "X1A" = "pink", "X1B" = "hotpink", "X2A" ="turquoise", "X2B" = "cyan")) +
  xlab("")

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Plots/qd_diversity/Shannon_Unthresholded.pdf", device = cairo_pdf, width = 3, height = 5)

```



#Plot untrhesholded Simpson diversity:
```{r}

p <- ggplot(data = simpson_diversity) +
  geom_bar(aes(x = Source, y = value, fill = Source), stat = "identity", width = 0.5) +
  scale_x_discrete(
    limits = c("Pre1","X1A","X1B","Pre2","X2A","X2B"),
    labels = c("Pre1" = "Pre-Treatment, Pool 1", "X1A" = "Post-Treatment, Pool 1, Replicate A", "X1B" = "Post-Treatment, Pool 1, Replicate B", "X2B" = "Post-Treatment, Pool 1, Replicate B", "Pre2" = "Pre-Treatment, Pool 2", "X2A" = "Post-Treatment, Pool 2, Replicate A", "X2B" = "Post-Treatment, Pool 2, Replicate B")
  ) +
  scale_y_continuous(
    name = "Simpson Diversity Index",
    expand = expansion(mult = c(0,0.05))
  ) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_text(family = font_type, size = font_size, color = "black", angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", margin = margin(t = 0, r = 0.05, b = 0, l = 0)),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.position = "none"
  ) +
  scale_fill_manual(values = c("Pre1" = "red", "Pre2" = "blue", "X1A" = "pink", "X1B" = "hotpink", "X2A" ="turquoise", "X2B" = "cyan")) +
  xlab("")

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Plots/qd_diversity/Simpson_Unthresholded.pdf", device = cairo_pdf, width = 3, height = 5)

```

#Generate data considering thresholded data:
```{r, include=FALSE}

gDNA_counts_thresholded <- gDNA_counts[rowSums(gDNA_counts >= read_threshold)>0,]

gDNA_proportion_thresholded <- sweep(gDNA_counts_thresholded, 2, colSums(gDNA_counts), "/")

```

#Get diversity metrics for thresholded reads:
```{r, include=FALSE}

number_of_unique_barcodes_thresholded <- colSums(gDNA_proportion_thresholded != 0)
number_of_unique_barcodes_thresholded <- melt(number_of_unique_barcodes_thresholded)
number_of_unique_barcodes_thresholded$Source <- rownames(number_of_unique_barcodes_thresholded)

shannon_diversity_thresholded <- sapply(gDNA_proportion_thresholded, function(x) vegan::diversity(x, index = "shannon"))
shannon_diversity_thresholded <- melt(shannon_diversity_thresholded)
shannon_diversity_thresholded$Source <- rownames(shannon_diversity_thresholded)

simpson_diversity_thresholded <- sapply(gDNA_proportion_thresholded, function(x) vegan::diversity(x, index = "simpson"))
simpson_diversity_thresholded <- melt(simpson_diversity_thresholded)
simpson_diversity_thresholded$Source <- rownames(simpson_diversity_thresholded)

```


#Plot thesholded number of unique barcodes:
```{r}

p <- ggplot(data = number_of_unique_barcodes_thresholded) +
  geom_bar(aes(x = Source, y = value, fill = Source), stat = "identity", width = 0.5) +
  scale_x_discrete(
    limits = c("Pre1","X1A","X1B","Pre2","X2A","X2B"),
    labels = c("Pre1" = "Pre-Treatment, Pool 1", "X1A" = "Post-Treatment, Pool 1, Replicate A", "X1B" = "Post-Treatment, Pool 1, Replicate B", "X2B" = "Post-Treatment, Pool 1, Replicate B", "Pre2" = "Pre-Treatment, Pool 2", "X2A" = "Post-Treatment, Pool 2, Replicate A", "X2B" = "Post-Treatment, Pool 2, Replicate B")
  ) +
  scale_y_continuous(
    name = "# of Unique Lineage Barcodes",
    expand = expansion(mult = c(0,0.05))
  ) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_text(family = font_type, size = font_size, color = "black", angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", margin = margin(t = 0, r = 0.05, b = 0, l = 0)),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.position = "none"
  ) +
  scale_fill_manual(values = c("Pre1" = "red", "Pre2" = "blue", "X1A" = "pink", "X1B" = "hotpink", "X2A" ="turquoise", "X2B" = "cyan")) +
  xlab("")

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Plots/qd_diversity/Number_of_Barcodes_Thresholded.pdf", device = cairo_pdf, width = 3, height = 5)

```


#Plot trhesholded Shannon diversity:
```{r}

p <- ggplot(data = shannon_diversity_thresholded) +
  geom_bar(aes(x = Source, y = value, fill = Source), stat = "identity", width = 0.5) +
  scale_x_discrete(
    limits = c("Pre1","X1A","X1B","Pre2","X2A","X2B"),
    labels = c("Pre1" = "Pre-Treatment, Pool 1", "X1A" = "Post-Treatment, Pool 1, Replicate A", "X1B" = "Post-Treatment, Pool 1, Replicate B", "X2B" = "Post-Treatment, Pool 1, Replicate B", "Pre2" = "Pre-Treatment, Pool 2", "X2A" = "Post-Treatment, Pool 2, Replicate A", "X2B" = "Post-Treatment, Pool 2, Replicate B")
  ) +
  scale_y_continuous(
    name = "Shannon Diversity Index",
    expand = expansion(mult = c(0,0.05))
  ) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_text(family = font_type, size = font_size, color = "black", angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", margin = margin(t = 0, r = 0.05, b = 0, l = 0)),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.position = "none"
  ) +
  scale_fill_manual(values = c("Pre1" = "red", "Pre2" = "blue", "X1A" = "pink", "X1B" = "hotpink", "X2A" ="turquoise", "X2B" = "cyan")) +
  xlab("")

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Plots/qd_diversity/Shannon_Thresholded.pdf", device = cairo_pdf, width = 3, height = 5)

```


#Plot trhesholded Simpson diversity:
```{r}

p <- ggplot(data = simpson_diversity_thresholded) +
  geom_bar(aes(x = Source, y = value, fill = Source), stat = "identity", width = 0.5) +
  scale_x_discrete(
    limits = c("Pre1","X1A","X1B","Pre2","X2A","X2B"),
    labels = c("Pre1" = "Pre-Treatment, Pool 1", "X1A" = "Post-Treatment, Pool 1, Replicate A", "X1B" = "Post-Treatment, Pool 1, Replicate B", "X2B" = "Post-Treatment, Pool 1, Replicate B", "Pre2" = "Pre-Treatment, Pool 2", "X2A" = "Post-Treatment, Pool 2, Replicate A", "X2B" = "Post-Treatment, Pool 2, Replicate B")
  ) +
  scale_y_continuous(
    name = "Simpson Diversity Index",
    expand = expansion(mult = c(0,0.05))
  ) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_text(family = font_type, size = font_size, color = "black", angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", margin = margin(t = 0, r = 0.05, b = 0, l = 0)),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.position = "none"
  ) +
  scale_fill_manual(values = c("Pre1" = "red", "Pre2" = "blue", "X1A" = "pink", "X1B" = "hotpink", "X2A" ="turquoise", "X2B" = "cyan")) +
  xlab("")

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Plots/qd_diversity/Simpson_Thresholded.pdf", device = cairo_pdf, width = 3, height = 5)

```



--------------------#Plot diversity metrics separately for each pool with thresholded gDNA---------------------------


#Plot thesholded number of unique barcodes, pool 1:
```{r}

p <- ggplot(data = number_of_unique_barcodes_thresholded) +
  geom_bar(aes(x = Source, y = value, fill = Source), stat = "identity", width = 0.5) +
  scale_x_discrete(
    limits = c("Pre1","X1A","X1B"),
    labels = c("Pre1" = "Pre-Treatment", "X1A" = "Post-Treatment, Rep. A", "X1B" = "Post-Treatment, Rep. B")
  ) +
  scale_y_continuous(
    name = "# of Unique Lineage Barcodes",
    expand = expansion(mult = c(0,0.05))
  ) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_text(family = font_type, size = font_size, color = "black", angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", margin = margin(t = 0, r = 0.05, b = 0, l = 0)),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.position = "none"
  ) +
  scale_fill_manual(values = c("Pre1" = "red", "X1A" = "pink", "X1B" = "hotpink")) +
  xlab("")

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Plots/qd_diversity/Number_of_Barcodes_Thresholded_Pool1.pdf", device = cairo_pdf, width = 3, height = 5)

```
