---
title: "Add_SCENIC_Data"
output: html_document
date: "2025-10-02"
---


```{r setup, include=FALSE}

#Run only in chunk:
knitr::opts_knit$set(root.dir = '/Users/robertvandervelde/Library/CloudStorage/GoogleDrive-robertvandervelde@sydshafferlab.com/My Drive/Vander_Velde_Share/Experiments/RV143')

getwd()

```


#Load Libraries:
```{r, include=FALSE}

rm(list = ls())
library(Seurat)
library(tidyverse)
library(broom)
library(Matrix)
library(reshape2)
library(dplyr)

library(egg)

library(Cairo)

library(ComplexHeatmap)

set.seed(1)

```

#Set font size and type:
```{r, include=FALSE}

font_size <- 8
font_type <- "Helvetica Neue"
point_size <- 0.01
jitter_width <- 0.2
point_stroke_weight <- 0.05
box_line_width <- 0.2
nudge_size <- 0.25

panel_dimensions <- c(0.4,1)

```

#Load strict barcode requirements seurat object:
```{r, include=FALSE}

seurat_strict_barcode_requirements <- readRDS("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Create_Seurat_Objects/seurat_strict_barcode_requirements")

```

#Use only transduced cells:
```{r, include=FALSE}

Idents(seurat_strict_barcode_requirements) <- "orig.ident"

seurat_strict_barcode_requirements_transduced_only <- subset(seurat_strict_barcode_requirements, idents = "Enriched")

rm(seurat_strict_barcode_requirements)

```

#Load SCENIC data:
```{r, include=FALSE}

auc_output <- read.csv("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/SCENIC_20251027/Output_Data/AUC_output.csv", check.names = FALSE , row.names = 1)


```

#Add SCENIC data as assay:
```{r, include=FALSE}

seurat_strict_barcode_requirements_transduced_only[["SCENIC_AUC"]] <- CreateAssayObject(data = t(as.matrix(auc_output)))

```


#Save data frame with AUC:
```{r, include=FALSE}

AUC_df <- data.frame(GetAssayData(seurat_strict_barcode_requirements_transduced_only, assay = "SCENIC_AUC"))

#Transpose:
AUC_df_transpose <- as.data.frame(t(AUC_df))

AUC_df_transpose$cell <- rownames(AUC_df_transpose)

#Melt dataframe:
melted_AUC_df <- melt(AUC_df_transpose, variable.name = "Transcription_Factor", id.vars = "cell")

#Restore "-" in cell names in order for them to match:
melted_AUC_df$cell <- lapply(melted_AUC_df$cell, function(x) gsub("\\.", "-", x))

#Get lineage assignment:
lineage_assignment <-  seurat_strict_barcode_requirements_transduced_only$Mini_Library_Assignment

#Add lineage assignment to dataframe:
melted_AUC_df$MiniLibrary_Assignment <- lineage_assignment[match(melted_AUC_df$cell, names(lineage_assignment))]

#Get list of cells with information on if the lineage barcode was enriched:
enriched_list <- seurat_strict_barcode_requirements_transduced_only$has_targeted_lineage_barcode

#Add enrichment status using matching cell barcodes:
melted_AUC_df$Enriched <- enriched_list[match(melted_AUC_df$cell, names(enriched_list))]

#Remove old dataframes:
rm(AUC_df, AUC_df_transpose, enriched_list)

Idents(seurat_strict_barcode_requirements_transduced_only) <- "has_targeted_lineage_barcode"

```

#Run grouped ANOVA:
```{r, include=FALSE}

ttest_results <- melted_AUC_df %>%
  group_by(Transcription_Factor) %>%
  do(tidy(t.test(value ~ Enriched, data = .))) %>%
  ungroup()

ttest_results$p.value.adjusted <- p.adjust(ttest_results$p.value, method = "BH")

sig_BH <- ttest_results[ttest_results$p.value.adjusted < 0.05,]

```

#Run find markers on SCENIC AUCs:
```{r, include=FALSE}

markers_targeted_vs_not <- FindMarkers(seurat_strict_barcode_requirements_transduced_only, ident.1 = "Targeted_Lineage_Barcode", ident.2 = "Not_Targeted_Lineage_Barcode", assay = "SCENIC_AUC")

```

#View grn:
```{r, include=FALSE}

grn_output <- read.table("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/SCENIC_20251027/Output_Data/GRN_output.tsv", header = TRUE)

ctx_output <- read.csv("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/SCENIC_20251027/Output_Data/CTX_output.csv", 
                       skip = 1,  # Skip row 1 (the "Enrichment" labels)
                       header = TRUE)

#Replace column names of CTX for TF and MotifID:
colnames(ctx_output)[1:2] <- c("TF", "MotifID")

#Remove ID row for CTX:
ctx_output <- ctx_output[2:dim(ctx_output)[1],]


```

#Create an outlier column for the AUC dataframe, to allow plotting as jitter with boxplot:
```{r, include=FALSE}

melted_AUC_df <- melted_AUC_df %>%
  group_by(Enriched, Transcription_Factor) %>%
  mutate(
    Q1 = quantile(value, 0.25, na.rm = TRUE),
    Q3 = quantile(value, 0.75, na.rm = TRUE),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR,
    is_outlier = value < lower_bound | value > upper_bound
  ) %>%
  ungroup() %>%
  select(-Q1, -Q3, -IQR, -lower_bound, -upper_bound) %>%
  as.data.frame()

```

#Plot FOSL1(+) as a simple boxplot:
```{r}

p <- ggplot(melted_AUC_df[melted_AUC_df$Transcription_Factor == "FOSL1(+)",],
       aes(x = factor(Enriched, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = value,
           color = factor(Enriched))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(melted_AUC_df[melted_AUC_df$Transcription_Factor == "FOSL1(+)",], is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("AUC") +
  ggtitle(paste("FOSL1(+)\nAdj. p-value:\n", format(markers_targeted_vs_not["FOSL1(+)",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Add_SCENIC_Data/FOSL1.pdf", device = cairo_pdf, width = 2, height = 3)
  
```


#Plot FOSL2(+) as a simple boxplot:
```{r}

p <- ggplot(melted_AUC_df[melted_AUC_df$Transcription_Factor == "FOSL2(+)",],
       aes(x = factor(Enriched, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = value,
           color = factor(Enriched))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(melted_AUC_df[melted_AUC_df$Transcription_Factor == "FOSL2(+)",], is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("AUC") +
  ggtitle(paste("FOSL2(+)\nAdj. p-value:\n", format(markers_targeted_vs_not["FOSL2(+)",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Add_SCENIC_Data/FOSL2.pdf", device = cairo_pdf, width = 2, height = 3)
  
```



#Plot FOS(+) as a simple boxplot:
```{r}

p <- ggplot(melted_AUC_df[melted_AUC_df$Transcription_Factor == "FOS(+)",],
       aes(x = factor(Enriched, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = value,
           color = factor(Enriched))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(melted_AUC_df[melted_AUC_df$Transcription_Factor == "FOS(+)",], is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("AUC") +
  ggtitle(paste("FOS(+)\nAdj. p-value:\n", format(markers_targeted_vs_not["FOS(+)",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Add_SCENIC_Data/FOS.pdf", device = cairo_pdf, width = 2, height = 3)
  
```

#Plot FOSB(+) as a simple boxplot:
```{r}

p <- ggplot(melted_AUC_df[melted_AUC_df$Transcription_Factor == "FOSB(+)",],
       aes(x = factor(Enriched, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = value,
           color = factor(Enriched))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(melted_AUC_df[melted_AUC_df$Transcription_Factor == "FOSB(+)",], is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("AUC") +
  ggtitle(paste("FOSB(+)\nAdj. p-value:\n", format(markers_targeted_vs_not["FOSB(+)",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Add_SCENIC_Data/FOSB.pdf", device = cairo_pdf, width = 2, height = 3)
  
```


#Plot JUN(+) as a simple boxplot:
```{r}

p <- ggplot(melted_AUC_df[melted_AUC_df$Transcription_Factor == "JUN(+)",],
       aes(x = factor(Enriched, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = value,
           color = factor(Enriched))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(melted_AUC_df[melted_AUC_df$Transcription_Factor == "JUN(+)",], is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("AUC") +
  ggtitle(paste("JUN(+)\nAdj. p-value:\n", format(markers_targeted_vs_not["JUN(+)",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Add_SCENIC_Data/JUN.pdf", device = cairo_pdf, width = 2, height = 3)
  
```


#Plot JUNB(+) as a simple boxplot:
```{r}

p <- ggplot(melted_AUC_df[melted_AUC_df$Transcription_Factor == "JUNB(+)",],
       aes(x = factor(Enriched, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = value,
           color = factor(Enriched))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(melted_AUC_df[melted_AUC_df$Transcription_Factor == "JUNB(+)",], is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("AUC") +
  ggtitle(paste("JUNB(+)\nAdj. p-value:\n", format(markers_targeted_vs_not["JUNB(+)",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Add_SCENIC_Data/JUNB.pdf", device = cairo_pdf, width = 2, height = 3)
  
```


#Plot JUND(+) as a simple boxplot:
```{r}

p <- ggplot(melted_AUC_df[melted_AUC_df$Transcription_Factor == "JUND(+)",],
       aes(x = factor(Enriched, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = value,
           color = factor(Enriched))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(melted_AUC_df[melted_AUC_df$Transcription_Factor == "JUND(+)",], is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("AUC") +
  ggtitle(paste("JUND(+)\nAdj. p-value:\n", format(markers_targeted_vs_not["JUND(+)",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Add_SCENIC_Data/JUND.pdf", device = cairo_pdf, width = 2, height = 3)
  
```


#Plot NFKB1(+) as a simple boxplot:
```{r}

p <- ggplot(melted_AUC_df[melted_AUC_df$Transcription_Factor == "NFKB1(+)",],
       aes(x = factor(Enriched, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = value,
           color = factor(Enriched))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(melted_AUC_df[melted_AUC_df$Transcription_Factor == "NFKB1(+)",], is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("AUC") +
  ggtitle(paste("NFKB1(+)\nAdj. p-value:\n", format(markers_targeted_vs_not["NFKB1(+)",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Add_SCENIC_Data/NFKB1.pdf", device = cairo_pdf, width = 2, height = 3)
  
```


#Plot NFKB2(+) as a simple boxplot:
```{r}

p <- ggplot(melted_AUC_df[melted_AUC_df$Transcription_Factor == "NFKB2(+)",],
       aes(x = factor(Enriched, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = value,
           color = factor(Enriched))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(melted_AUC_df[melted_AUC_df$Transcription_Factor == "NFKB2(+)",], is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("AUC") +
  ggtitle(paste("NFKB2(+)\nAdj. p-value:\n", format(markers_targeted_vs_not["NFKB2(+)",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Add_SCENIC_Data/NFKB2.pdf", device = cairo_pdf, width = 2, height = 3)
  
```


#Plot CREB3(+) as a simple boxplot:
```{r}

p <- ggplot(melted_AUC_df[melted_AUC_df$Transcription_Factor == "CREB3(+)",],
       aes(x = factor(Enriched, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = value,
           color = factor(Enriched))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(melted_AUC_df[melted_AUC_df$Transcription_Factor == "CREB3(+)",], is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("AUC") +
  ggtitle(paste("CREB3(+)\nAdj. p-value:\n", format(markers_targeted_vs_not["CREB3(+)",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Add_SCENIC_Data/CREB3.pdf", device = cairo_pdf, width = 2, height = 3)
  
```


#Plot CREB3L2(+) as a simple boxplot:
```{r}

p <- ggplot(melted_AUC_df[melted_AUC_df$Transcription_Factor == "CREB3L2(+)",],
       aes(x = factor(Enriched, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = value,
           color = factor(Enriched))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(melted_AUC_df[melted_AUC_df$Transcription_Factor == "CREB3L2(+)",], is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("AUC") +
  ggtitle(paste("CREB3L2(+)\nAdj. p-value:\n", format(markers_targeted_vs_not["CREB3L2(+)",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Add_SCENIC_Data/CREB3L2.pdf", device = cairo_pdf, width = 2, height = 3)
  
```


#Plot RUNX3(+) as a simple boxplot:
```{r}

p <- ggplot(melted_AUC_df[melted_AUC_df$Transcription_Factor == "RUNX3(+)",],
       aes(x = factor(Enriched, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = value,
           color = factor(Enriched))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(melted_AUC_df[melted_AUC_df$Transcription_Factor == "RUNX3(+)",], is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("AUC") +
  ggtitle(paste("RUNX3(+)\nAdj. p-value:\n", format(markers_targeted_vs_not["RUNX3(+)",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Add_SCENIC_Data/RUNX3.pdf", device = cairo_pdf, width = 2, height = 3)
  
```


#Get significant transcription factors for heatmap plotting:
```{r, include=FALSE}

sig_tfs_list <- rownames(markers_targeted_vs_not[markers_targeted_vs_not$p_val_adj < 0.05,])

sig_tfs_df <- markers_targeted_vs_not[sig_tfs_list, "avg_log2FC", drop = FALSE]

sig_tfs_matrix <- as.matrix(t(sig_tfs_df))

```

#Plot heatmap of significant transcription factor:
```{r}

CairoPDF("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Add_SCENIC_Data/Sig_TF_Heatmap.pdf", width = 7, height = 6)

ComplexHeatmap::Heatmap(sig_tfs_matrix,
                        show_column_dend = FALSE,
                        column_names_side = "top",
                        show_row_names = FALSE,
                        heatmap_legend_param = list(title = "Avg. Log2(Enriched / Non-Enriched)",
                                                    labels_gp = gpar(fontsize = font_size-0.5, fontfamily = font_type),
                                                    title_gp = gpar(fontsize = font_size-0.5, fontfamily = font_type)),
                        column_names_gp = gpar(fontsize = font_size-0.5, fontfamily = font_type),
                        #Set height using the number of columns:
                        width =  unit(0.15*ncol(sig_tfs_matrix), "in"),
                        height = unit(0.5, "in")
)

dev.off()

```
#Get info on motifs for :
```{r, include=FALSE}

#Get ctx just for significant TFs:
ctx_sig_tfs <- ctx_output %>%
  filter(TF %in% gsub("\\(\\+\\)|\\(\\-\\)", "", sig_tfs_list))

unique_MotifID_sig_TFs <- unique(ctx_sig_tfs$MotifID)

ctx_sig_tfs[ctx_sig_tfs$MotifID == "tfdimers__MD00558",]

ctx_sig_tfs[ctx_sig_tfs$TF == "FOS",]

ctx_sig_tfs[ctx_sig_tfs$TF == "FOSL1",]

ctx_sig_tfs[ctx_sig_tfs$TF == "FOSB",]

ctx_sig_tfs[ctx_sig_tfs$MotifID == "transfac_pro__M07666",]






```



#Try making a UMAP from SCENIC AUC:
```{r, include=FALSE}

DefaultAssay(seurat_strict_barcode_requirements_transduced_only) <- "SCENIC_AUC"

#Get assay data to rename rows:
scenic_auc_matrix <- LayerData(seurat_strict_barcode_requirements_transduced_only, 
                                assay = "SCENIC_AUC", 
                                layer = "data")

# Clean up the rownames
rownames(scenic_auc_matrix) <- gsub("\\(\\+\\)", "-Pos", rownames(scenic_auc_matrix))
rownames(scenic_auc_matrix) <- gsub("\\(\\-\\)", "-Neg", rownames(scenic_auc_matrix))

#Replace SCENIC AUC values with cleaned names:
seurat_strict_barcode_requirements_transduced_only[["SCENIC_AUC"]] <- CreateAssayObject(counts = scenic_auc_matrix)

seurat_strict_barcode_requirements_transduced_only <- ScaleData(
  seurat_strict_barcode_requirements_transduced_only,
  features = rownames(seurat_strict_barcode_requirements_transduced_only[["SCENIC_AUC"]])
)

seurat_strict_barcode_requirements_transduced_only <- FindVariableFeatures(seurat_strict_barcode_requirements_transduced_only)

seurat_strict_barcode_requirements_transduced_only <- RunPCA(
  seurat_strict_barcode_requirements_transduced_only,
  features = rownames(seurat_strict_barcode_requirements_transduced_only[["SCENIC_AUC"]]),
  reduction.name = "pca_SCENIC",
  reduction.key = "PC_SCENIC_"
)


seurat_strict_barcode_requirements_transduced_only <- FindNeighbors(
  seurat_strict_barcode_requirements_transduced_only, 
  reduction = "pca_SCENIC",
  dims = 1:10,
  graph.name = "SCENIC_snn"
)

seurat_strict_barcode_requirements_transduced_only <- FindClusters(
  seurat_strict_barcode_requirements_transduced_only, 
  graph.name = "SCENIC_snn",
  resolution = 0.5
)

seurat_strict_barcode_requirements_transduced_only <- RunUMAP(
  seurat_strict_barcode_requirements_transduced_only,
  reduction = "pca_SCENIC",
  dims = 1:10,
  reduction.name = "umap_SCENIC",
  reduction.key = "UMAPSCENIC_"
)


```

#View dim plots:
```{r}

DimPlot(seurat_strict_barcode_requirements_transduced_only, 
              reduction = "pca_SCENIC", 
              group.by = "seurat_clusters")

DimPlot(seurat_strict_barcode_requirements_transduced_only, 
              reduction = "umap_SCENIC", 
              group.by = "seurat_clusters")

DimPlot(seurat_strict_barcode_requirements_transduced_only, 
              reduction = "pca_SCENIC", 
              group.by = "has_targeted_lineage_barcode")

DimPlot(seurat_strict_barcode_requirements_transduced_only, 
              reduction = "umap_SCENIC", 
              group.by = "has_targeted_lineage_barcode")

```




