---
title: "Evenness_In_Single_Cell_Barcode_Data"
output: html_document
date: "2025-10-21"
---



```{r setup, include=FALSE}

#Run only in chunk:
knitr::opts_knit$set(root.dir = '/Users/robertvandervelde/Library/CloudStorage/GoogleDrive-robertvandervelde@sydshafferlab.com/My Drive/Vander_Velde_Share/Experiments/RV143')

getwd()

```


#Load Libraries:
```{r, include=FALSE}

rm(list = ls())
library(Seurat)
library(tidyverse)
library(Matrix)

library(egg)

library(Cairo)

set.seed(1)

```


#Set font size and type:
```{r, include=FALSE}

font_size <- 8
font_type <- "Helvetica Neue"
point_size <- 0.01
jitter_width <- 0.2
point_stroke_weight <- 0.05
box_line_width <- 0.2
nudge_size <- 0.25

panel_dimensions <- c(0.4,1)

```

#Load strict barcode requirements seurat object:
```{r, include=FALSE}

seurat_strict_barcode_requirements <- readRDS("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Create_Seurat_Objects/seurat_strict_barcode_requirements")

```

#Assign lineage barcodes based on barcode with most counts in that cell:
```{r, include=FALSE}

lineage_barcode_count_matrix <- GetAssayData(seurat_strict_barcode_requirements, assay = "Lineage_Assay", layer = "counts")

seurat_strict_barcode_requirements$Lineage_Barcode_w_Max_Count <- apply(lineage_barcode_count_matrix, 2, function(x) {
  rownames(lineage_barcode_count_matrix)[which.max(x)]
})

rm(lineage_barcode_count_matrix)


```

#Create tables 
```{r, include=FALSE}

Idents(seurat_strict_barcode_requirements) <- "orig.ident"

table_untransduced <- table(table(seurat_strict_barcode_requirements$Lineage_Barcode_w_Max_Count[seurat_strict_barcode_requirements$orig.ident == "Untransduced"]))

table_transduced <- table(table(seurat_strict_barcode_requirements$Lineage_Barcode_w_Max_Count[seurat_strict_barcode_requirements$orig.ident == "Enriched"]))

table_percent_untransduced <- as.numeric(names(table_untransduced))*100 / sum(as.numeric(names(table_untransduced))*table_untransduced)

table_percent_transduced <- as.numeric(names(table_transduced))*100 / sum(as.numeric(names(table_transduced))*table_transduced)

plot(table_untransduced)

plot(table_transduced)

#Creat dataframes for tables:
table_percent_untransduced_df <- data.frame("percent_cells" = table_percent_untransduced,
                                            "number_of_barcodes" = as.numeric(table_untransduced))

table_percent_transduced_df <- data.frame("percent_cells" = table_percent_transduced,
                                          "number_of_barcodes" = as.numeric(table_transduced))

```

GGATCGTAAGGTGGGAGCGT

GG A TCGTAA GGTGGGAGCGT

TCGAGCGTGCTAGCGTTCCC

#Hypothesis, GGGAAGGTGGGTGGGAGCGT in enrichment library is enriching for GGATCGTAAGGTGGGAGCGT

GGATCGTAAGGTGGGAGCGT is more common in Pre2: 2.576259e-05 proportion in GGGAAGGTGGGTGGGAGCGT and 3.612825e-05 in GGATCGTAAGGTGGGAGCGT

TTATCGGATCAACTTTGTATAGAAAAGTTGTACTGTGGTATCACTCGTC GG G AAGGTG GGTGGGAGCGT AGGCTCAGCTAAGGTGCAATGGTAGGCGTGTACGGTGGGAGGCCTATATAA

TTATCGGATCAACTTTGTATAGAAAAGTTGTACTGTGGTATCACTCGTC TCGAGCGTGCTAGCGTTCCC AGGCTCAGCTAAGGTGCAATGGTAGGCGTGTACGGTGGGAGGCCTATATAA


```{r}


p <- ggplot(data = table_percent_untransduced_df) +
  geom_segment(aes(x = percent_cells, xend = percent_cells, 
                   y = 0, yend = number_of_barcodes), 
               linewidth = 0.1) +
  scale_x_continuous(
    name = "% of Cells per Lineage Barcode",
    limits = c(0,0.25),
    expand = expansion(mult = c(0,0.05))
  ) +
  scale_y_continuous(
    name = "Number of Lineage Barcodes with % of Cells",
    expand = expansion(mult = c(0,0.05))
  ) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_text(family = font_type, size = font_size, color = "black", angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", margin = margin(t = 0, r = 0.05, b = 0, l = 0)),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.position = "none"
  )

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))


```




