---
title: "enricheR"
output: html_document
date: "2025-10-20"
---

```{r setup, include=FALSE}

#Run only in chunk:
knitr::opts_knit$set(root.dir = '/Users/robertvandervelde/Library/CloudStorage/GoogleDrive-robertvandervelde@sydshafferlab.com/My Drive/Vander_Velde_Share/Experiments/RV143')

getwd()

```

#Load Libraries:
```{r, include=FALSE}

rm(list = ls())

library(GSEABase)
library(enrichplot)
library(org.Hs.eg.db)
library(clusterProfiler)
library(tidyverse)
library(msigdbr)
library(Cairo)
library(ggplot2)
library(egg)

```

#Set font size and type:
```{r, include=FALSE}

font_size <- 8
font_type <- "Helvetica Neue"

panel_dimensions <- c(0.8,0.6)

```

#Set thresholds:
```{r, include=FALSE}

pct_threshold <- 0.05

```

#Load differential expression data from find markers:
```{r, include=FALSE}

getwd()

markers_untranduced_vs_transduced <- read.csv("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Compare_Differential_Expression/markers_untranduced_vs_transduced.csv", row.names = 1)

markers_targeted_vs_not <- read.csv("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Compare_Differential_Expression/markers_targeted_vs_not.csv", row.names = 1)

```

#Subset to genes that are significant with adjusted p-value and detected in >= 5% for targeted vs untargeted:
```{r, include=FALSE}

markers_targeted_vs_not_thresholded <- markers_targeted_vs_not[((markers_targeted_vs_not$pct.1 >= pct_threshold) & (markers_targeted_vs_not$pct.2 >= pct_threshold)) & (markers_targeted_vs_not$p_val_adj < 0.05),]

#split into negative and positive fold change:
markers_targeted_vs_not_thresholded_positive <- markers_targeted_vs_not_thresholded[markers_targeted_vs_not_thresholded$avg_log2FC > 0, ]

markers_targeted_vs_not_thresholded_negative <- markers_targeted_vs_not_thresholded[markers_targeted_vs_not_thresholded$avg_log2FC < 0, ]

```

#Get gene sets for human:
```{r, include=FALSE}

gene_sets_Hallmarks <- msigdbr(species = "Homo sapiens", collection = "H") %>%
  dplyr::select(gs_name, source_gene)

go_bp_sets <- msigdbr(species = "Homo sapiens", collection = "C5", subcollection = "GO:BP") %>%
  select(gs_name, gene_symbol)

```

#Run geneset analysis:
```{r, include=FALSE}

positive_GSEA_targeted_Hallmarks <- enricher(row.names(markers_targeted_vs_not_thresholded_positive), TERM2GENE=gene_sets_Hallmarks)

negative_GSEA_targeted_Hallmarks <- enricher(row.names(markers_targeted_vs_not_thresholded_negative), TERM2GENE=gene_sets_Hallmarks)




positive_GSEA_targeted_bp <- enricher(row.names(markers_targeted_vs_not_thresholded_positive), TERM2GENE=go_bp_sets)

negative_GSEA_targeted_bp <- enricher(row.names(markers_targeted_vs_not_thresholded_negative), TERM2GENE=go_bp_sets)




#Check for melanocyte related programs:
positive_GSEA_targeted_bp@result %>%
  filter(grepl("MEL", ID, ignore.case = TRUE))

negative_GSEA_targeted_bp@result %>%
  filter(grepl("MEL", ID, ignore.case = TRUE))


```

#Plot bubble plot:
```{r}

#Get min and max gene ratios for plotting:
positive_GSEA_targeted_Hallmarks@result$GeneRatio_numeric <- 
  sapply(strsplit(as.character(positive_GSEA_targeted_Hallmarks@result$GeneRatio), "/"), 
         function(x) as.numeric(x[1]) / as.numeric(x[2]))

#Get min and max ratios for plotting:
min_significant_fraction_hallmarks <- min(positive_GSEA_targeted_Hallmarks@result[positive_GSEA_targeted_Hallmarks@result$p.adjust < 0.05,]$GeneRatio_numeric)

max_significant_fraction_hallmarks <- max(positive_GSEA_targeted_Hallmarks@result[positive_GSEA_targeted_Hallmarks@result$p.adjust < 0.05,]$GeneRatio_numeric)

#Get min and max gene numbers for plotting:
min_significant_gene_number_hallmarks <- min(positive_GSEA_targeted_Hallmarks@result[positive_GSEA_targeted_Hallmarks@result$p.adjust < 0.05,]$Count)

max_significant_gene_number_hallmarks <- max(positive_GSEA_targeted_Hallmarks@result[positive_GSEA_targeted_Hallmarks@result$p.adjust < 0.05,]$Count)

p <- dotplot(positive_GSEA_targeted_Hallmarks) +
  scale_x_continuous(
    breaks = c(min_significant_fraction_hallmarks, max_significant_fraction_hallmarks),
    limits = c(min_significant_fraction_hallmarks,
               max_significant_fraction_hallmarks),
    expand = c(0.005, 0.005)
  ) + 
  scale_size_continuous(
    breaks = c(min_significant_gene_number_hallmarks, 
               max_significant_gene_number_hallmarks)
  ) +
  theme(
    axis.text.x = element_text(size = font_size, family = font_type, angle = 90, vjust = 0.5),
    axis.text.y = element_text(size = font_size, family = font_type),
    axis.title.x = element_text(size = font_size, family = font_type),
    axis.title.y = element_text(size = font_size, family = font_type),
    legend.text = element_text(size = font_size, family = font_type),
    legend.title = element_text(size = font_size, family = font_type),
    plot.title = element_text(size = font_size, family = font_type)
  )

p

#Change panel size:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))


ggsave(p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Gene_Set_Analysis/dotplot_hallmarks.pdf",
       device = cairo_pdf,
       width = 6, 
       height = 3.5)

```

#Get positive_GSEA_targeted_Hallmarks as dataframe for plotting:
```{r}

positive_GSEA_targeted_Hallmarks_df <- as.data.frame(positive_GSEA_targeted_Hallmarks)

positive_GSEA_targeted_Hallmarks_df <- positive_GSEA_targeted_Hallmarks_df[order(positive_GSEA_targeted_Hallmarks_df$GeneRatio_numeric),]

```

#Plot bar plot:
```{r}

ggplot(positive_GSEA_targeted_Hallmarks_df, aes(x = GeneRatio_numeric, y = reorder(ID, GeneRatio_numeric), fill = p.adjust)) +
  geom_col(width = 0.5) +
  theme_bw() +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +
  ylab("") +
  xlab("Gene Ratio") +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.text = element_text(family = font_type, size = font_size, color = "black")
  )

ggsave(filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Gene_Set_Analysis/barplot_pos_hallmarks.pdf",
       device = cairo_pdf,
       width = 5, 
       height = 1.2)

```



#Plot bar plot, vertical p-value as labels:
```{r}

ggplot(positive_GSEA_targeted_Hallmarks_df, aes(x = reorder(ID, GeneRatio_numeric), y = GeneRatio_numeric)) +
  geom_col(width = 0.5) +
  geom_text(aes(label = sprintf("Adj. p-value =\n %.2e", p.adjust)), 
            vjust = -0.5, 
            family = font_type, 
            size = font_size / .pt,
            color = "black") +
  theme_bw() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  ylab("Gene Ratio") +
  xlab("Gene Set") +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text.x = element_text(family = font_type, size = font_size, color = "black", angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.text = element_text(family = font_type, size = font_size, color = "black")
  )

ggsave(filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Gene_Set_Analysis/barplot_pos_hallmarks_pvalues.pdf",
       device = cairo_pdf,
       width = 2, 
       height = 5)

```




