---
title: "Plot_UMAPs"
output: html_document
date: "2025-09-29"
---


```{r setup, include=FALSE}

#Run only in chunk:
knitr::opts_knit$set(root.dir = '/Users/robertvandervelde/Library/CloudStorage/GoogleDrive-robertvandervelde@sydshafferlab.com/My Drive/Vander_Velde_Share/Experiments/RV143')

getwd()

```


#Load Libraries:
```{r, include=FALSE}

rm(list = ls())
library(Seurat)
library(ggplot2)
library(scattermore)
library(dplyr)

library(ggh4x)

library(Polychrome)

getwd()

set.seed(1)

```

#Set variable:
```{r, include=FALSE}

font_size <- 8
font_type <- "Helvetica Neue"

```

#Load strict barcode requirements seurat object:
```{r, include=FALSE}

seurat_strict_barcode_requirements <- readRDS("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Create_Seurat_Objects/seurat_strict_barcode_requirements")

```

#Get umap data into a dataframe:
```{r, include=FALSE}

umap_embeddings <- as.data.frame(Embeddings(seurat_strict_barcode_requirements, reduction = "umap"))

#Add lineage assignment to umap_embeddings 
umap_embeddings <- cbind(umap_embeddings, seurat_strict_barcode_requirements$Mini_Library_Assignment)

#Add orignal identity to umap_embeddings 
umap_embeddings <- cbind(umap_embeddings, seurat_strict_barcode_requirements$orig.ident)

colnames(umap_embeddings) <- c("umap_1", "umap_2", "Mini_Library_Assignment", "Protocol")

```

#Get UMAP ranges for consistent point sizes and plot sizes:
```{r, include=FALSE}

#Get ranges for consistent plotting:

#Get UMAP ranges for consistent point sizes and plot sizes:
umap_range_x <- max(umap_embeddings$umap_1) - min(umap_embeddings$umap_1)
umap_range_y <- max(umap_embeddings$umap_2) - min(umap_embeddings$umap_2)

# Define max_range for equal scaling
max_range <- max(umap_range_x, umap_range_y)

x_center <- mean(range(umap_embeddings$umap_1))
y_center <- mean(range(umap_embeddings$umap_2))

x_limits <- c(x_center - max_range/2, x_center + max_range/2)
y_limits <- c(y_center - max_range/2, y_center + max_range/2)

#Set panel size:
panel_size <- unit(1.5, "in")

```

#Plot umap with unstransduced and transduced:
```{r}

ggplot(umap_embeddings, aes(x = umap_1, y = umap_2, color = Protocol)) +
  scattermore::geom_scattermore(pointsize = 2, 
                                pixels = c(umap_range_x*50, umap_range_x*50)) +
  labs(x = "UMAP 1", 
       y = "UMAP 2") +
  scale_x_continuous(limits = x_limits) +
  scale_y_continuous(limits = y_limits) +
  scale_color_discrete(breaks = c("Untransduced","Enriched"), 
                      labels = c("Untransduced" = "Enrichment Library Negative",
                                "Enriched" = "Enrichment Library Positive")) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.position = "none"
  ) +
  coord_fixed() +  # This enforces 1:1 visual scale
  force_panelsizes(rows = panel_size, cols = panel_size)

ggsave("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_UMAPs/Enrichment_Negative_Positive.pdf", plot = last_plot(), width = 3, height = 3, device = cairo_pdf)


```

#Create a column for enriched vs non-enriched lineages:
```{r, include=FALSE}

#Get if cell has targeted lineage barcode:
umap_embeddings$has_targeted_lineage_barcode <- seurat_strict_barcode_requirements$has_targeted_lineage_barcode

#Create dataframe with just transduced:
transduced_only_umap_embeddings <- umap_embeddings[umap_embeddings$Protocol == "Enriched",]

#Get dataframe with enriched lineage barcodes only for plotting to the front:
enriched_umap_embeddings <- transduced_only_umap_embeddings[transduced_only_umap_embeddings$has_targeted_lineage_barcode == "Targeted_Lineage_Barcode",]

#Get dataframe without enriched lineage barcodes for plotting in back:
non_enriched_umap_embeddings <- transduced_only_umap_embeddings[transduced_only_umap_embeddings$has_targeted_lineage_barcode == "Not_Targeted_Lineage_Barcode",]


```


#Set panel size for UMAPs without axes:
```{r, include=FALSE}

#Set panel size:
panel_size <- unit(2, "in")

```

#Plot all lineage barcodes together:
```{r}

ggplot() +
  scattermore::geom_scattermore(data = non_enriched_umap_embeddings, 
                                aes(x = umap_1, y = umap_2, color = has_targeted_lineage_barcode),
                                pointsize = 2, 
                                pixels = c(umap_range_x*50, umap_range_x*50)) +
  scattermore::geom_scattermore(data = enriched_umap_embeddings, 
                                aes(x = umap_1, y = umap_2, color = Mini_Library_Assignment),
                                pointsize = 5, 
                                pixels = c(umap_range_x*50, umap_range_x*50)) +
  scale_color_manual(values = c("Non-Enriched" = "grey",
                                "TCGAGCGTGCTAGCGTTCCC" = "black",
                                "CGGTCGTATCCTTGGAGGTA" = "#654522",
                                "TGAACGGTCCTTAGGTGGCT" = "#8db600",
                                "TGTTCGTTCGGATGCTAGGT" = "#a1caf1",
                                "GCCAGCATGCATTGGTGCCA" = "#be0032",
                                "TGTAGGGAACATAGTATGAT" = "#c2b280",
                                "GGGTACGACGAACCGTACGA" = "#008856",
                                "GGATCGCAGCTAGGTAGGCT" = "#e68fac",
                                "GCTATGTTGCAAGCATTGAT" = "#0067a5",
                                "ACTATCATTGCAAGCTTCGA" = "#2b3d26",
                                "CGAAGGGAGCTAGCTTACGT" = "#604e97",
                                "TCGAGGTTAGATTCTATCTT" = "#f6a600",
                                "GGGTTCAAACTATCCATGAT" = "#b3446c",
                                "GGGAAGGTGGGTGGGAGCGT" = "#dcd300",
                                "GGTTGGTAAGTTTGTTAGGA" = "#882d17"
                                )) + 
  labs(x = "UMAP 1", y = "UMAP 2") +
  scale_x_continuous(limits = x_limits) +
  scale_y_continuous(limits = y_limits) +
  theme_void() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", angle = 90),
    legend.position = "none"
  ) +
  coord_fixed(ratio = 1) +
  force_panelsizes(rows = panel_size, cols = panel_size)

ggsave("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_UMAPs/Targeted_Barcodes_Together.pdf", 
       plot = last_plot(), width = 2, height = 2, device = cairo_pdf)

rep(0.06666667,15)

pdf("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_UMAPs/Pie_Point.pdf")

pie(rep(0.06666667,15),
    labels = "",
    col = c("black", "#654522", "#8db600", "#a1caf1","#be0032","#c2b280","#008856", "#e68fac","#0067a5","#2b3d26","#604e97","#f6a600","#b3446c", "#dcd300","#882d17" )) 

dev.off()



```

#Plot all enriched lineage barcodes together, without non-enriched barcodes:
```{r}

p <- ggplot() +
  #Add a rasterized data to ensure alignment:
  scattermore::geom_scattermore(data = umap_embeddings, 
                                aes(x = umap_1, y = umap_2),
                                pointsize = 0,
                                pixels = c(umap_range_x*50, umap_range_x*50),
                                alpha = 0) +
  geom_point(data = enriched_umap_embeddings, 
                                aes(x = umap_1, y = umap_2, color = Mini_Library_Assignment),
                                size = 1,
                                stroke = 0) +
  scale_color_manual(values = c("Non-Enriched" = "grey",
                                "TCGAGCGTGCTAGCGTTCCC" = "#604e97",
                                "CGGTCGTATCCTTGGAGGTA" = "#654522",
                                "TGAACGGTCCTTAGGTGGCT" = "#8db600",
                                "TGTTCGTTCGGATGCTAGGT" = "#a1caf1",
                                "GCCAGCATGCATTGGTGCCA" = "#be0032",
                                "TGTAGGGAACATAGTATGAT" = "#c2b280",
                                "GGGTACGACGAACCGTACGA" = "#008856",
                                "GGATCGCAGCTAGGTAGGCT" = "#e68fac",
                                "GCTATGTTGCAAGCATTGAT" = "#0067a5",
                                "ACTATCATTGCAAGCTTCGA" = "#2b3d26",
                                "CGAAGGGAGCTAGCTTACGT" = "grey50",
                                "TCGAGGTTAGATTCTATCTT" = "#f6a600",
                                "GGGTTCAAACTATCCATGAT" = "#b3446c",
                                "GGGAAGGTGGGTGGGAGCGT" = "#dcd300",
                                "GGTTGGTAAGTTTGTTAGGA" = "#882d17"
                                )) + 
  labs(x = "UMAP 1", y = "UMAP 2") +
  scale_x_continuous(limits = x_limits) +
  scale_y_continuous(limits = y_limits) +
  theme_void() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", angle = 90),
    legend.position = "none"
  ) +
  coord_fixed(ratio = 1)

p

p <- egg::set_panel_size(p, width = panel_size, height = panel_size)

ggsave(p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_UMAPs/Targeted_Barcodes_Together_No_NonEnriched.pdf", width = 4, height = 4, device = cairo_pdf)


```

#Plot UMAP of source without background:
```{r}

p <- ggplot(umap_embeddings, aes(x = umap_1, y = umap_2, color = Protocol)) +
  scattermore::geom_scattermore(pointsize = 2, 
                                pixels = c(umap_range_x*50, umap_range_x*50)) + 
  scale_color_discrete(breaks = c("Untransduced","Enriched"), 
                      labels = c("Untransduced" = "Enrichment Library Negative",
                                "Enriched" = "Enrichment Library Positive")) +
  labs(x = "UMAP 1", y = "UMAP 2") +
  scale_x_continuous(limits = x_limits) +
  scale_y_continuous(limits = y_limits) +
  theme_void() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", angle = 90)
  ) +
  coord_fixed(ratio = 1) +
  force_panelsizes(rows = panel_size, cols = panel_size)

p

p <- egg::set_panel_size(p, width = panel_size, height = panel_size)

ggsave(p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_UMAPs/Source_UMAP.pdf", width = 4, height = 4, device = cairo_pdf)


```


#Plot UMAP of targeted barcodes in enrichment library negative:
```{r}

p <- ggplot() +
  #Add a rasterized data to ensure alignment:
  scattermore::geom_scattermore(data = umap_embeddings, 
                                aes(x = umap_1, y = umap_2),
                                pointsize = 0,
                                pixels = c(umap_range_x*50, umap_range_x*50),
                                alpha = 0) +
  geom_point(data = umap_embeddings[(umap_embeddings$Protocol == "Untransduced") & (umap_embeddings$has_targeted_lineage_barcode == "Targeted_Lineage_Barcode"),], 
                                aes(x = umap_1, y = umap_2, color = Mini_Library_Assignment),
                                size = 1,
                                stroke = 0) +
  scale_color_manual(values = c("Non-Enriched" = "grey",
                                "TCGAGCGTGCTAGCGTTCCC" = "#604e97",
                                "CGGTCGTATCCTTGGAGGTA" = "#654522",
                                "TGAACGGTCCTTAGGTGGCT" = "#8db600",
                                "TGTTCGTTCGGATGCTAGGT" = "#a1caf1",
                                "GCCAGCATGCATTGGTGCCA" = "#be0032",
                                "TGTAGGGAACATAGTATGAT" = "#c2b280",
                                "GGGTACGACGAACCGTACGA" = "#008856",
                                "GGATCGCAGCTAGGTAGGCT" = "#e68fac",
                                "GCTATGTTGCAAGCATTGAT" = "#0067a5",
                                "ACTATCATTGCAAGCTTCGA" = "#2b3d26",
                                "CGAAGGGAGCTAGCTTACGT" = "grey50",
                                "TCGAGGTTAGATTCTATCTT" = "#f6a600",
                                "GGGTTCAAACTATCCATGAT" = "#b3446c",
                                "GGGAAGGTGGGTGGGAGCGT" = "#dcd300",
                                "GGTTGGTAAGTTTGTTAGGA" = "#882d17"
                                )) + 
  labs(x = "UMAP 1", y = "UMAP 2") +
  scale_x_continuous(limits = x_limits) +
  scale_y_continuous(limits = y_limits) +
  theme_void() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", angle = 90),
    legend.position = "none"
  ) +
  coord_fixed(ratio = 1)

p

p <- egg::set_panel_size(p, width = panel_size, height = panel_size)


ggsave(p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_UMAPs/Targeted_Barcodes_No_Enrichment_Library.pdf", width = 4, height = 4, device = cairo_pdf)

```





