---
title: "UMAP_Analysis"
output: html_document
date: "2025-08-18"
---

```{r setup, include=FALSE}

#Run only in chunk:
knitr::opts_knit$set(root.dir = '/Users/robertvandervelde/Library/CloudStorage/GoogleDrive-robertvandervelde@sydshafferlab.com/My Drive/Vander_Velde_Share/Experiments/RV143')

getwd()

```


#Load Libraries:
```{r, include=FALSE}

rm(list = ls())
library(Seurat)
library(ggplot2)
library(scattermore)
library(dplyr)

library(ggh4x)

library(Polychrome)

library(egg)

getwd()

set.seed(1)

```

#Set variable:
```{r, include=FALSE}

font_size <- 8
font_type <- "Helvetica Neue"

```

#Load strict barcode requirements seurat object:
```{r, include=FALSE}

seurat_strict_barcode_requirements <- readRDS("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Create_Seurat_Objects/seurat_strict_barcode_requirements")

```

#Load cell-cycle genes:
```{r, include=FALSE}

s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

```

#Add cell cycle scores:
```{r, include=FALSE}

seurat_strict_barcode_requirements <- CellCycleScoring(seurat_strict_barcode_requirements, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

Idents(seurat_strict_barcode_requirements) <- seurat_strict_barcode_requirements$Phase

DimPlot(seurat_strict_barcode_requirements, reduction = "umap")

head(seurat_strict_barcode_requirements$Phase)

```

#Regress out cell cycle:
```{r, include=FALSE}

seurat_strict_barcode_requirements <- ScaleData(seurat_strict_barcode_requirements, assay = "RNA", vars.to.regress = c("S.Score", "G2M.Score"), features = rownames(seurat_strict_barcode_requirements))

#View current assays:
Assays(seurat_strict_barcode_requirements)
Layers(seurat_strict_barcode_requirements)


```

#Generate dimension reduction without cell cycle:
```{r}

seurat_strict_barcode_requirements <- FindVariableFeatures(seurat_strict_barcode_requirements)

seurat_strict_barcode_requirements <- RunPCA(seurat_strict_barcode_requirements, features = VariableFeatures(object = seurat_strict_barcode_requirements))

seurat_strict_barcode_requirements <- FindNeighbors(seurat_strict_barcode_requirements, dims = 1:10)
seurat_strict_barcode_requirements <- FindClusters(seurat_strict_barcode_requirements, resolution = 0.5)

seurat_strict_barcode_requirements <- RunUMAP(seurat_strict_barcode_requirements, dims = 1:10)

#Plot cell cycle score with cell-cycle regressed out:
DimPlot(seurat_strict_barcode_requirements, reduction = "umap")

```

#Save seurat object with cell cycle regressed out:

```{r, include=FALSE}

saveRDS(seurat_strict_barcode_requirements,"RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Create_Seurat_Objects/seurat_strict_barcode_requirements_cell_cycle_regression")

```

#Load cell cycle regressed out seurat object:

```{r, include=FALSE}

seurat_strict_barcode_requirements <- readRDS("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Create_Seurat_Objects/seurat_strict_barcode_requirements_cell_cycle_regression")

```

#Get umap data into a dataframe:
```{r, include=FALSE}

umap_embeddings <- as.data.frame(Embeddings(seurat_strict_barcode_requirements, reduction = "umap"))

#Add lineage assignment to umap_embeddings 
umap_embeddings <- cbind(umap_embeddings, seurat_strict_barcode_requirements$Mini_Library_Assignment)

#Add orignal identity to umap_embeddings 
umap_embeddings <- cbind(umap_embeddings, seurat_strict_barcode_requirements$orig.ident)

colnames(umap_embeddings) <- c("umap_1", "umap_2", "Mini_Library_Assignment", "Protocol")

```

#Get UMAP ranges for consistent point sizes and plot sizes:
```{r, include=FALSE}

#Get ranges for consistent plotting:

#Get UMAP ranges for consistent point sizes and plot sizes:
umap_range_x <- max(umap_embeddings$umap_1) - min(umap_embeddings$umap_1)
umap_range_y <- max(umap_embeddings$umap_2) - min(umap_embeddings$umap_2)

# Define max_range for equal scaling
max_range <- max(umap_range_x, umap_range_y)

x_center <- mean(range(umap_embeddings$umap_1))
y_center <- mean(range(umap_embeddings$umap_2))

x_limits <- c(x_center - max_range/2, x_center + max_range/2)
y_limits <- c(y_center - max_range/2, y_center + max_range/2)

#Set panel size:
panel_size <- unit(1.5, "in")

```

#Create a column for enriched vs non-enriched lineages:
```{r, include=FALSE}

#Get if cell has targeted lineage barcode:
umap_embeddings$has_targeted_lineage_barcode <- seurat_strict_barcode_requirements$has_targeted_lineage_barcode

#Create dataframe with just transduced:
transduced_only_umap_embeddings <- umap_embeddings[umap_embeddings$Protocol == "Enriched",]

#Get dataframe with enriched lineage barcodes only for plotting to the front:
enriched_umap_embeddings <- transduced_only_umap_embeddings[transduced_only_umap_embeddings$has_targeted_lineage_barcode == "Targeted_Lineage_Barcode",]

#Get dataframe without enriched lineage barcodes for plotting in back:
non_enriched_umap_embeddings <- transduced_only_umap_embeddings[transduced_only_umap_embeddings$has_targeted_lineage_barcode == "Not_Targeted_Lineage_Barcode",]

```


#Set panel size for UMAPs without axes:
```{r, include=FALSE}

#Set panel size:
panel_size <- unit(2, "in")

```

#Plot all lineage barcodes together:
```{r}

ggplot() +
  scattermore::geom_scattermore(data = non_enriched_umap_embeddings, 
                                aes(x = umap_1, y = umap_2, color = has_targeted_lineage_barcode),
                                pointsize = 2, 
                                pixels = c(umap_range_x*50, umap_range_x*50)) +
  scattermore::geom_scattermore(data = enriched_umap_embeddings, 
                                aes(x = umap_1, y = umap_2, color = Mini_Library_Assignment),
                                pointsize = 5, 
                                pixels = c(umap_range_x*50, umap_range_x*50)) +
  scale_color_manual(values = c("Non-Enriched" = "grey",
                                "TCGAGCGTGCTAGCGTTCCC" = "black",
                                "CGGTCGTATCCTTGGAGGTA" = "#654522",
                                "TGAACGGTCCTTAGGTGGCT" = "#8db600",
                                "TGTTCGTTCGGATGCTAGGT" = "#a1caf1",
                                "GCCAGCATGCATTGGTGCCA" = "#be0032",
                                "TGTAGGGAACATAGTATGAT" = "#c2b280",
                                "GGGTACGACGAACCGTACGA" = "#008856",
                                "GGATCGCAGCTAGGTAGGCT" = "#e68fac",
                                "GCTATGTTGCAAGCATTGAT" = "#0067a5",
                                "ACTATCATTGCAAGCTTCGA" = "#2b3d26",
                                "CGAAGGGAGCTAGCTTACGT" = "#604e97",
                                "TCGAGGTTAGATTCTATCTT" = "#f6a600",
                                "GGGTTCAAACTATCCATGAT" = "#b3446c",
                                "GGGAAGGTGGGTGGGAGCGT" = "#dcd300",
                                "GGTTGGTAAGTTTGTTAGGA" = "#882d17"
                                )) + 
  labs(x = "UMAP 1", y = "UMAP 2") +
  scale_x_continuous(limits = x_limits) +
  scale_y_continuous(limits = y_limits) +
  theme_void() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", angle = 90),
    legend.position = "none"
  ) +
  coord_fixed(ratio = 1) +
  force_panelsizes(rows = panel_size, cols = panel_size)

ggsave("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Cell_Cycle_UMAPs/Targeted_Barcodes_Together.pdf", 
       plot = last_plot(), width = 2, height = 2, device = cairo_pdf)


```


#Plot all enriched lineage barcodes together, without non-enriched barcodes:
```{r}

p <- ggplot() +
  #Add a rasterized data to ensure alignment:
  scattermore::geom_scattermore(data = umap_embeddings, 
                                aes(x = umap_1, y = umap_2),
                                pointsize = 0,
                                pixels = c(umap_range_x*50, umap_range_x*50),
                                alpha = 0) +
  geom_point(data = enriched_umap_embeddings, 
                                aes(x = umap_1, y = umap_2, color = Mini_Library_Assignment),
                                size = 1,
                                stroke = 0) +
  scale_color_manual(values = c("Non-Enriched" = "grey",
                                "TCGAGCGTGCTAGCGTTCCC" = "#604e97",
                                "CGGTCGTATCCTTGGAGGTA" = "#654522",
                                "TGAACGGTCCTTAGGTGGCT" = "#8db600",
                                "TGTTCGTTCGGATGCTAGGT" = "#a1caf1",
                                "GCCAGCATGCATTGGTGCCA" = "#be0032",
                                "TGTAGGGAACATAGTATGAT" = "#c2b280",
                                "GGGTACGACGAACCGTACGA" = "#008856",
                                "GGATCGCAGCTAGGTAGGCT" = "#e68fac",
                                "GCTATGTTGCAAGCATTGAT" = "#0067a5",
                                "ACTATCATTGCAAGCTTCGA" = "#2b3d26",
                                "CGAAGGGAGCTAGCTTACGT" = "grey50",
                                "TCGAGGTTAGATTCTATCTT" = "#f6a600",
                                "GGGTTCAAACTATCCATGAT" = "#b3446c",
                                "GGGAAGGTGGGTGGGAGCGT" = "#dcd300",
                                "GGTTGGTAAGTTTGTTAGGA" = "#882d17"
                                )) + 
  labs(x = "UMAP 1", y = "UMAP 2") +
  scale_x_continuous(limits = x_limits) +
  scale_y_continuous(limits = y_limits) +
  theme_void() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", angle = 90),
    legend.position = "none"
  ) +
  coord_fixed(ratio = 1)

p

p <- egg::set_panel_size(p, width = panel_size, height = panel_size)

ggsave(p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Cell_Cycle_UMAPs/Targeted_Barcodes_Together_No_NonEnriched.pdf", width = 4, height = 4, device = cairo_pdf)


```


#Plot UMAP of source without background:
```{r}

p <- ggplot(umap_embeddings, aes(x = umap_1, y = umap_2, color = Protocol)) +
  scattermore::geom_scattermore(pointsize = 2, 
                                pixels = c(umap_range_x*50, umap_range_x*50)) + 
  scale_color_discrete(breaks = c("Untransduced","Enriched"), 
                      labels = c("Untransduced" = "Enrichment Library Negative",
                                "Enriched" = "Enrichment Library Positive")) +
  labs(x = "UMAP 1", y = "UMAP 2") +
  scale_x_continuous(limits = x_limits) +
  scale_y_continuous(limits = y_limits) +
  theme_void() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", angle = 90)
  ) +
  coord_fixed(ratio = 1)

p

p <- egg::set_panel_size(p, width = panel_size, height = panel_size)



ggsave(p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Cell_Cycle_UMAPs/Source_UMAP.pdf", width = 4, height = 4, device = cairo_pdf)


```


#Plot UMAP of targeted barcodes in enrichment library negative:
```{r}

p <- ggplot() +
  #Add a rasterized data to ensure alignment:
  scattermore::geom_scattermore(data = umap_embeddings, 
                                aes(x = umap_1, y = umap_2),
                                pointsize = 0,
                                pixels = c(umap_range_x*50, umap_range_x*50),
                                alpha = 0) +
  geom_point(data = umap_embeddings[(umap_embeddings$Protocol == "Untransduced") & (umap_embeddings$has_targeted_lineage_barcode == "Targeted_Lineage_Barcode"),], 
                                aes(x = umap_1, y = umap_2, color = Mini_Library_Assignment),
                                size = 1,
                                stroke = 0) +
  scale_color_manual(values = c("Non-Enriched" = "grey",
                                "TCGAGCGTGCTAGCGTTCCC" = "#604e97",
                                "CGGTCGTATCCTTGGAGGTA" = "#654522",
                                "TGAACGGTCCTTAGGTGGCT" = "#8db600",
                                "TGTTCGTTCGGATGCTAGGT" = "#a1caf1",
                                "GCCAGCATGCATTGGTGCCA" = "#be0032",
                                "TGTAGGGAACATAGTATGAT" = "#c2b280",
                                "GGGTACGACGAACCGTACGA" = "#008856",
                                "GGATCGCAGCTAGGTAGGCT" = "#e68fac",
                                "GCTATGTTGCAAGCATTGAT" = "#0067a5",
                                "ACTATCATTGCAAGCTTCGA" = "#2b3d26",
                                "CGAAGGGAGCTAGCTTACGT" = "grey50",
                                "TCGAGGTTAGATTCTATCTT" = "#f6a600",
                                "GGGTTCAAACTATCCATGAT" = "#b3446c",
                                "GGGAAGGTGGGTGGGAGCGT" = "#dcd300",
                                "GGTTGGTAAGTTTGTTAGGA" = "#882d17"
                                )) + 
  labs(x = "UMAP 1", y = "UMAP 2") +
  scale_x_continuous(limits = x_limits) +
  scale_y_continuous(limits = y_limits) +
  theme_void() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.x = element_text(family = font_type, size = font_size, color = "black"),
    axis.title.y = element_text(family = font_type, size = font_size, color = "black", angle = 90),
    legend.position = "none"
  ) +
  coord_fixed(ratio = 1)

p

p <- egg::set_panel_size(p, width = panel_size, height = panel_size)



ggsave(p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Cell_Cycle_UMAPs/Targeted_Barcodes_No_Enrichment_Library.pdf", width = 4, height = 4, device = cairo_pdf)


```
