---
title: "Compare_Differential_Expression"
output: html_document
date: "2025-06-26"
---


```{r setup, include=FALSE}

#Run only in chunk:
knitr::opts_knit$set(root.dir = '/Users/robertvandervelde/Library/CloudStorage/GoogleDrive-robertvandervelde@sydshafferlab.com/My Drive/Vander_Velde_Share/Experiments/RV143')

getwd()

```


#Load Libraries:
```{r, include=FALSE}

rm(list = ls())
library(Seurat)
library(tidyverse)
library(Matrix)
library(dplyr)
library(ggrepel)
library(ggrastr)
library(scattermore)

library(egg)

set.seed(1)

```

#Set variable:
```{r, include=FALSE}

font_size <- 8
font_type <- "Helvetica Neue"

panel_dimensions <- c(1.5,1)

                   

```



#Load strict barcode requirements seurat object:
```{r, include=FALSE}

seurat_strict_barcode_requirements <- readRDS("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Create_Seurat_Objects/seurat_strict_barcode_requirements")

```

#Create Seurat object without enriched barcodes to properly compare untransduced and transduced samples:
```{r, include=FALSE}

seurat_strict_barcode_requirements_no_enriched <- subset(x=seurat_strict_barcode_requirements, subset = has_targeted_lineage_barcode == "Not_Targeted_Lineage_Barcode")

```

#Get markers for untransduced vs enriched (transduced):
```{r, include=FALSE}

Idents(seurat_strict_barcode_requirements_no_enriched) <- "orig.ident"

#Find markers, only use pct threshold so allow comparisons to other comparison:
markers_untranduced_vs_transduced <- FindMarkers(
  seurat_strict_barcode_requirements_no_enriched, 
  ident.1 = "Enriched", 
  ident.2 = "Untransduced",
  logfc.threshold = 0
)

#Remove extra seurat object:
rm(seurat_strict_barcode_requirements_no_enriched)

```

#Get transduced (enriched) sample only:
```{r, include=FALSE}


transduced_only <- subset(x=seurat_strict_barcode_requirements, subset = orig.ident == "Enriched")


```

```{r, include=FALSE}

#Change idents to be linage barcode assignment:
Idents(transduced_only) <- transduced_only$has_targeted_lineage_barcode

#Find markers, only use pct threshold so allow comparisons to other comparison:
markers_targeted_vs_not <- FindMarkers(
  transduced_only,
  ident.1 = "Targeted_Lineage_Barcode",
  ident.2 = "Not_Targeted_Lineage_Barcode",
  logfc.threshold = 0
)


#Remove extra seurat object:
rm(transduced_only)

```

#Save differential expression dataframes:
```{r, include=FALSE}

write.csv(markers_untranduced_vs_transduced, "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Compare_Differential_Expression/markers_untranduced_vs_transduced.csv")

write.csv(markers_targeted_vs_not, "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Compare_Differential_Expression/markers_targeted_vs_not.csv")

```

#Read differential expression dataframes:
```{r, include=FALSE}

getwd()

markers_untranduced_vs_transduced <- read.csv("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Compare_Differential_Expression/markers_untranduced_vs_transduced.csv", row.names = 1)

markers_targeted_vs_not <- read.csv("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Compare_Differential_Expression/markers_targeted_vs_not.csv", row.names = 1)

```

#Combine dataframes and add significance binary columns:
```{r, include=FALSE}

#Get genes that are in both datasets:
genes_in_both <- intersect(rownames(markers_untranduced_vs_transduced),rownames(markers_targeted_vs_not))

#Create a temporary dataframe for markers_untranduced_vs_transduced:
markers_untranduced_vs_transduced_temp <- markers_untranduced_vs_transduced

#Change markers_untranduced_vs_transduced colnames to add identifier:
colnames(markers_untranduced_vs_transduced_temp) <- paste0("untrans_vs_trans_", colnames(markers_untranduced_vs_transduced))

#Create a temporary dataframe for markers_targeted_vs_not:
markers_targeted_vs_not_temp <- markers_targeted_vs_not

#Change markers_targeted_vs_not colnames to add identifier:
colnames(markers_targeted_vs_not_temp) <- paste0("targeted_vs_not_", colnames(markers_targeted_vs_not))

#Combine the dataframes:
overall_find_markers <- merge(markers_untranduced_vs_transduced_temp, markers_targeted_vs_not_temp, by = "row.names")

#Remove temporary dataframes:
rm(markers_untranduced_vs_transduced_temp,markers_targeted_vs_not_temp)

#Make row names genes:
colnames(overall_find_markers)[1] <- "Genes"

#Create a column for significance threshold:
overall_find_markers$untrans_vs_trans_sig <- ifelse(overall_find_markers$untrans_vs_trans_p_val_adj < 0.05, TRUE, FALSE)

#Create a column for significance threshold:
overall_find_markers$targeted_vs_not_sig <- ifelse(overall_find_markers$targeted_vs_not_p_val_adj < 0.05, TRUE, FALSE)


overall_find_markers <- overall_find_markers %>%
  mutate(significance_category = case_when(
    !untrans_vs_trans_sig & !targeted_vs_not_sig ~ "Neither",
    untrans_vs_trans_sig & !targeted_vs_not_sig ~ "Untrans_vs_Trans",
    !untrans_vs_trans_sig & targeted_vs_not_sig ~ "Untargeted_vs_Targeted",
    untrans_vs_trans_sig & targeted_vs_not_sig ~ "Both"
  ))

#Change order of overall_find_markers for better plotting:
overall_find_markers <- overall_find_markers %>%
  arrange(match(significance_category, c("Neither", "Untrans_vs_Trans","Both", "Untargeted_vs_Targeted")))


```

#Plot the logFC
```{r}

#Create axis range for consistent axes:
axis_range <- range(c(overall_find_markers$untrans_vs_trans_avg_log2FC, 
                     overall_find_markers$targeted_vs_not_avg_log2FC), na.rm = TRUE)


ggplot(overall_find_markers, aes(x = untrans_vs_trans_avg_log2FC, y = targeted_vs_not_avg_log2FC, color = significance_category)) +
  rasterize(geom_point(size = 0.1), dpi = 1200) +
  labs(x = "Avg. log2(Enrichment Library +/Enrichment Library -)", 
       y = "Avg. log2(Enriched/Non-Enriched)",
       color = "Significant In...") +
  theme_bw() +
  coord_equal(xlim = axis_range, ylim = axis_range) +
  scale_color_manual(values = c("Neither" = "gray60",
                               "Untrans_vs_Trans" = "blue",
                               "Untargeted_vs_Targeted" = "red",
                               "Both" = "#D67DFF"),
                     breaks = c("Neither", "Untrans_vs_Trans", "Untargeted_vs_Targeted", "Both"),
                     labels = c("Neither", "Enrichment Library - vs Enrichment Library +", "Non-Enriched vs Enriched", "Both")) +
  #Make the legend points larger:
  guides(color = guide_legend(override.aes = list(size = 1))) +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.title = element_text(family = font_type, size = font_size, color = "black", margin = margin(b=0.1, unit = "cm")),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.spacing.y = unit(0, "cm"),
    legend.key.size = unit(0.2, "cm")
  )


ggsave("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Compare_Differential_Expression/Volcanoish.pdf", plot = last_plot(), device = cairo_pdf, width = 4, height = 3)


```



#Plot the logFC with Rectangle:
```{r}

#Create axis range for consistent axes:
axis_range <- range(c(overall_find_markers$untrans_vs_trans_avg_log2FC, 
                     overall_find_markers$targeted_vs_not_avg_log2FC), na.rm = TRUE)


ggplot(overall_find_markers, aes(x = untrans_vs_trans_avg_log2FC, y = targeted_vs_not_avg_log2FC, color = significance_category)) +
  geom_point(size = 0.5) +
  geom_rect(aes(xmin = -2, xmax = 2, ymin = -2, ymax = 2), 
            fill = NA, color = "black", linewidth = 0.5) +
  labs(x = "Avg. log2(Enrichment Library +/Enrichment Library -)", 
       y = "Avg. log2(Enriched/Non-Enriched)",
       color = "Significant In...") +
  theme_bw() +
  coord_equal(xlim = axis_range, ylim = axis_range) +
  scale_color_manual(values = c("Neither" = "gray60",
                               "Untrans_vs_Trans" = "blue",
                               "Untargeted_vs_Targeted" = "red",
                               "Both" = "#D67DFF"),
                     breaks = c("Neither", "Untrans_vs_Trans", "Untargeted_vs_Targeted", "Both"),
                     labels = c("Neither", "Enrichment Library - vs Enrichment Library +", "Non-Enriched vs Enriched", "Both")) +
  #Make the legend points larger:
  guides(color = guide_legend(override.aes = list(size = 3))) +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.title = element_text(family = font_type, size = font_size, color = "black", margin = margin(b=0.1, unit = "cm")),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.spacing.y = unit(0, "cm"),
    legend.key.size = unit(0.2, "cm")
  )


ggsave("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Compare_Differential_Expression/Volcanoish_with_Rectangle.pdf", plot = last_plot(), device = cairo_pdf, width = 10, height = 6)


```


#Get genes to label:

```{r, include=FALSE}

num_genes_to_label <- 4

x_top4 <- overall_find_markers %>%
  arrange(desc(untrans_vs_trans_avg_log2FC)) %>%
  slice_head(n = num_genes_to_label)

x_bottom4 <- overall_find_markers %>%
  arrange(untrans_vs_trans_avg_log2FC) %>%
  slice_head(n = num_genes_to_label)

# For y-axis (targeted_vs_not_avg_log2FC)
y_top4 <- overall_find_markers %>%
  arrange(desc(targeted_vs_not_avg_log2FC)) %>%
  slice_head(n = num_genes_to_label)

y_bottom4 <- overall_find_markers %>%
  arrange(targeted_vs_not_avg_log2FC) %>%
  slice_head(n = num_genes_to_label)

# Combine all genes to label
genes_to_label <- bind_rows(x_top4, x_bottom4, y_top4, y_bottom4) %>%
  distinct()


```



#Make plot with some genes labeled:
```{r}

#Create axis range for consistent axes:
axis_range <- range(c(overall_find_markers$untrans_vs_trans_avg_log2FC, 
                     overall_find_markers$targeted_vs_not_avg_log2FC), na.rm = TRUE)


ggplot(overall_find_markers, aes(x = untrans_vs_trans_avg_log2FC, y = targeted_vs_not_avg_log2FC, color = significance_category)) +
  rasterize(geom_point(size = 0.1), dpi = 1200) +
  labs(x = "Avg. log2(Enrichment Library +/Enrichment Library -)", 
       y = "Avg. log2(Enriched/Non-Enriched)",
       color = "Significant In...") +
  theme_bw() +
  coord_equal(xlim = axis_range, ylim = axis_range) +
  scale_color_manual(values = c("Neither" = "gray60",
                               "Untrans_vs_Trans" = "blue",
                               "Untargeted_vs_Targeted" = "red",
                               "Both" = "#D67DFF"),
                     breaks = c("Neither", "Untrans_vs_Trans", "Untargeted_vs_Targeted", "Both"),
                     labels = c("Neither", "Enrichment Library - vs Enrichment Library +", "Non-Enriched vs Enriched", "Both")) +
  #Make the legend points larger:
  guides(color = guide_legend(override.aes = list(size = 1))) +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.title = element_text(family = font_type, size = font_size, color = "black", margin = margin(b=0.1, unit = "cm")),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.spacing.y = unit(0, "cm"),
    legend.key.size = unit(0.2, "cm")
  )


ggsave("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Compare_Differential_Expression/Volcanoish.pdf", plot = last_plot(), device = cairo_pdf, width = 4, height = 3)


```

#Get genes that either have pct >= 0.05 in both untransduced and transduced or in both targeted and untargeted:
```{r, include=FALSE}

pct_threshold <- 0.05

overall_find_markers_pct_threshold <- overall_find_markers[((overall_find_markers$untrans_vs_trans_pct.1 >= pct_threshold) & (overall_find_markers$untrans_vs_trans_pct.2 >= pct_threshold)) | ((overall_find_markers$targeted_vs_not_pct.1 >= pct_threshold) & (overall_find_markers$targeted_vs_not_pct.2 >= pct_threshold)),]

overall_find_markers[overall_find_markers$Genes == "LINC01198",]

overall_find_markers_pct_threshold[overall_find_markers_pct_threshold$Genes == "LINC01198",]

overall_find_markers_pct_threshold[overall_find_markers_pct_threshold$Genes == "SNED1",]

overall_find_markers_pct_threshold[overall_find_markers_pct_threshold$Genes == "EGFR",]


```



#Make a plot for pct threshold:
```{r}

#Create axis range for consistent axes:
axis_range <- c(-4,4)


ggplot(overall_find_markers_pct_threshold, aes(x = untrans_vs_trans_avg_log2FC, y = targeted_vs_not_avg_log2FC, color = significance_category)) +
  rasterize(geom_point(size = 0.01), dpi = 1200) +
  labs(x = "Avg. log2(Enrichment Library +/Enrichment Library -)", 
       y = "Avg. log2(Enriched/Non-Enriched)",
       color = "Significant In...") +
  theme_bw() +
  coord_equal(xlim = axis_range, ylim = axis_range) +
  scale_color_manual(values = c("Neither" = "gray60",
                               "Untrans_vs_Trans" = "blue",
                               "Untargeted_vs_Targeted" = "red",
                               "Both" = "#D67DFF"),
                     breaks = c("Neither", "Untrans_vs_Trans", "Untargeted_vs_Targeted", "Both"),
                     labels = c("Neither", "Enrichment Library - vs Enrichment Library +", "Non-Enriched vs Enriched", "Both")) +
  #Make the legend points larger:
  guides(color = guide_legend(override.aes = list(size = 1))) +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.title = element_text(family = font_type, size = font_size, color = "black", margin = margin(b=0.1, unit = "cm")),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.spacing.y = unit(0, "cm"),
    legend.key.size = unit(0.2, "cm")
  )


ggsave("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Compare_Differential_Expression/Volcanoish_pct_threshold.pdf", plot = last_plot(), device = cairo_pdf, width = 5, height = 2.5)


```

#Plot with top and bottom genes in Enriched/Non-Enriched:
```{r}

#Create axis range for consistent axes:
axis_range <- c(-4,4)

top_genes <- overall_find_markers_pct_threshold %>%
  arrange(desc(targeted_vs_not_avg_log2FC)) %>%
  head(3)

bottom_genes <- overall_find_markers_pct_threshold %>%
  arrange(targeted_vs_not_avg_log2FC) %>%
  head(3)

#Put genes together to label them:
genes_to_label <- bind_rows(top_genes, bottom_genes)

#Add exact positions for labelled genes:
genes_to_label <- genes_to_label %>%
  mutate(
    label_x = c(-1.5, 0.5, -1, 1, -1.7, 0),  # x positions for labels
    label_y = c(4, 4, 3.5, -3.7, -4, -3),  # y positions for labels
    hjust_val = c(1, 0, 1, 0, 1, 0),
    vjust_val = c(0.5, 0.5, 0.5, 0.5, 0.5, 0.5)
  )


ggplot(overall_find_markers_pct_threshold, aes(x = untrans_vs_trans_avg_log2FC, y = targeted_vs_not_avg_log2FC, color = significance_category)) +
  rasterize(geom_point(size = 0.1), dpi = 2400) +
  geom_segment(data = genes_to_label,
               aes(x = untrans_vs_trans_avg_log2FC, 
                   y = targeted_vs_not_avg_log2FC,
                   xend = label_x, 
                   yend = label_y),
               color = "black",
               linewidth = 0.3,
               inherit.aes = FALSE) +
  geom_text(data = genes_to_label,
            aes(x = label_x, y = label_y, label = Genes, hjust = hjust_val, vjust = vjust_val),
            size = font_size / .pt,
            color = "black",
            family = font_type,
            inherit.aes = FALSE) +
  labs(x = "Avg. log2(Enrichment Library +/Enrichment Library -)", 
       y = "Avg. log2(Enriched/Non-Enriched)",
       color = "Significant In...") +
  theme_bw() +
  coord_equal(xlim = axis_range, ylim = axis_range) +
  scale_color_manual(values = c("Neither" = "gray60",
                               "Untrans_vs_Trans" = "blue",
                               "Untargeted_vs_Targeted" = "red",
                               "Both" = "#D67DFF"),
                     breaks = c("Neither", "Untrans_vs_Trans", "Untargeted_vs_Targeted", "Both"),
                     labels = c("Neither", "Enrichment Library - vs Enrichment Library +", "Non-Enriched vs Enriched", "Both")) +
  #Make the legend points larger:
  guides(color = guide_legend(override.aes = list(size = 1))) +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.title = element_text(family = font_type, size = font_size, color = "black", margin = margin(b=0.1, unit = "cm")),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.spacing.y = unit(0, "cm"),
    legend.key.size = unit(0.2, "cm")
  )


ggsave("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Compare_Differential_Expression/Volcanoish_pct_threshold_labels.pdf", plot = last_plot(), device = cairo_pdf, width = 5, height = 2.5)


```

#Get -log10 of adjusted p-values for faster plotting:
```{r, include=FALSE}

overall_find_markers$untrans_vs_trans_p_val_adj_log10 <- -log10(overall_find_markers$untrans_vs_trans_p_val_adj)

overall_find_markers$targeted_vs_not_p_val_adj_log10 <- -log10(overall_find_markers$targeted_vs_not_p_val_adj)

```

#Get genes pct >= 0.05 in both untransduced and transduced or in both targeted and untargeted, separately, for separate volcano plots:
```{r, include=FALSE}

overall_find_markers_pct_threshold_transduced_vs_not <- overall_find_markers[((overall_find_markers$untrans_vs_trans_pct.1 >= pct_threshold) & (overall_find_markers$untrans_vs_trans_pct.2 >= pct_threshold)),]

overall_find_markers_pct_threshold_targeted_vs_not <- overall_find_markers[((overall_find_markers$targeted_vs_not_pct.1 >= pct_threshold) & (overall_find_markers$targeted_vs_not_pct.2 >= pct_threshold)),]

```


#Plot transduced vs not regular volcano plot:
```{r}


#Sort for plotting:
overall_find_markers_pct_threshold_transduced_vs_not <- overall_find_markers_pct_threshold_transduced_vs_not %>%
  arrange(desc(untrans_vs_trans_sig))

p <- ggplot(overall_find_markers_pct_threshold_transduced_vs_not, aes(x = untrans_vs_trans_avg_log2FC, y = untrans_vs_trans_p_val_adj_log10, color = untrans_vs_trans_sig)) +
  ggrastr::rasterize(geom_point(size = 0.5, stroke = 0), dpi = 2400) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.title = element_text(family = font_type, size = font_size, color = "black", margin = margin(b=0.1, unit = "cm")),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.spacing.y = unit(0, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.position = "none",
    panel.border = element_blank(),
  ) +
  scale_y_continuous(expand = c(0,0)) +
  scale_color_manual(values = c("FALSE" = "black",
                     "TRUE" = "blue"),
                     breaks = c("FALSE",
                                "TRUE")) +
  labs(x = "Avg. log2(Enrichment Library +/Enrichment Library -)", 
       y = "log10(Adj. p-value)")

p

p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Compare_Differential_Expression/Volcano_Untransduced_vs_Transduced.pdf", device = cairo_pdf, width = 4, height = 3)

```


#Plot transduced vs not regular volcano plot:
```{r}

#Sort for plotting:
overall_find_markers_pct_threshold_targeted_vs_not <- overall_find_markers_pct_threshold_targeted_vs_not %>%
  arrange(desc(targeted_vs_not_sig))

p <- ggplot(overall_find_markers_pct_threshold_targeted_vs_not, aes(x = targeted_vs_not_avg_log2FC, y = targeted_vs_not_p_val_adj_log10, color = targeted_vs_not_sig)) +
  ggrastr::rasterize(geom_point(size = 0.5, stroke = 0), dpi = 2400) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.title = element_text(family = font_type, size = font_size, color = "black", margin = margin(b=0.1, unit = "cm")),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    legend.spacing.y = unit(0, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.position = "none",
    panel.border = element_blank(),
  ) +
  scale_y_continuous(expand = c(0,0)) +
  scale_color_manual(values = c("FALSE" = "black",
                     "TRUE" = "red"),
                     breaks = c("FALSE",
                                "TRUE")) +
  labs(x = "Avg. log2(Enriched/Non-Enriched)", 
       y = "log10(Adj. p-value)")

p

p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Compare_Differential_Expression/Volcano_Targeted_vs_Not.pdf", device = cairo_pdf, width = 4, height = 3)

```


