---
title: "Create_Seurat_Objects"
output: html_document
date: "2025-05-23"
---


```{r setup, include=FALSE}

#Run only in chunk:
knitr::opts_knit$set(root.dir = '/Users/robertvandervelde/Library/CloudStorage/GoogleDrive-robertvandervelde@sydshafferlab.com/My Drive/Vander_Velde_Share/Experiments/RV143')

getwd()

```


#Load Libraries:
```{r, include=FALSE}

rm(list = ls())
library(Seurat)
library(tidyverse)
library(Matrix)

library(stringr)

library(pheatmap)

getwd()

set.seed(1)

```

#Set parameters:
```{r, include=FALSE}

lineage_barcode_umi_threshold <- 3

```

#Create custom functions:
```{r, include=FALSE}

#Create function that prints non-zero lineage barcodes in a cell:
print_non_zero_barcodes <- function(cell_name, data, diff_value) {
  if (is.na(cell_name)) {
    return()
  }
  
  cell_data <- data[, cell_name]
  non_zero <- cell_data[cell_data > 0]
  non_zero_sorted <- sort(non_zero, decreasing = TRUE)
  
  cat("  Cell:", cell_name, "\n")
  cat("  Barcodes:", paste(names(non_zero_sorted), "=", non_zero_sorted, collapse=", "), "\n\n")
}

```

#Load data from counts:
```{r, include=FALSE}

counts_untransduced <- Read10X("RV143_Comb_20250407_20250508_Seq/20250518_RV143_Enrichment_counts_untransduced/outs/raw_feature_bc_matrix")

counts_enriched <- Read10X("RV143_Comb_20250407_20250508_Seq/20250518_RV143_Enrichment_counts_enriched/outs/raw_feature_bc_matrix")

```


#Create Seurat objects:
```{r, include=FALSE}

seurat_object_untransduced <- CreateSeuratObject(counts = counts_untransduced, project = "Untransduced", min.cells = 3, min.features = 200)

seurat_object_enriched <- CreateSeuratObject(counts = counts_enriched, project = "Enriched", min.cells = 3, min.features = 200)

```

#Remove initial matrices to save space:
```{r, include=FALSE}

rm(counts_untransduced,counts_enriched)

```


#Plot features for filtering:
```{r}

seurat_object_untransduced[["percent.mt"]] <- PercentageFeatureSet(seurat_object_untransduced, pattern = "^MT-")
VlnPlot(seurat_object_untransduced, layer = "counts", features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

seurat_object_enriched[["percent.mt"]] <- PercentageFeatureSet(seurat_object_enriched, pattern = "^MT-")
VlnPlot(seurat_object_enriched, layer = "counts", features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

```

Filter cells in Seurat objects and remove original matrices:
```{r, include=FALSE}

seurat_object_untransduced_standard_filter <- subset(seurat_object_untransduced, subset = nFeature_RNA > 200 & nFeature_RNA < 9000 & nCount_RNA > 10000 & percent.mt < 10)

rm(seurat_object_untransduced)

seurat_object_enriched_standard_filter <- subset(seurat_object_enriched, subset = nFeature_RNA > 200 & nFeature_RNA < 9000 & nCount_RNA > 10000 & percent.mt < 10)

rm(seurat_object_enriched)

```


#Load barcode count dataframe:
```{r, include=FALSE}

barcode_data <- read.table("RV143_Comb_20250407_20250508_Seq/20250520_Unmapped/EB_output_modified.txt")

colnames(barcode_data) <- c("Cell_Barcode","UMI","Lineage_Barcode","StartSeq_Type","Sample")

```

#Get counts of each UMI per cell barcode, lineage barcode combination, this will be added to the seurat object as an assay:
```{r, include=FALSE}

#Do this for untransduced cells:

#Subset untransduced cells:
barcode_data_untransduced <- barcode_data[barcode_data$Sample == "untransduced_step1_unmapped.txt",]

#Do this for 
lineage_counts_untransduced <- barcode_data_untransduced %>%
  # Group by Cell_Barcode and Lineage_Barcode
  group_by(Lineage_Barcode, Cell_Barcode) %>%
  # Count unique UMIs
  summarize(count = n_distinct(UMI)) %>%
  # Ungroup
  ungroup() %>%
  # Pivot wider to create matrix-like format needed for Seurat
  pivot_wider(
    names_from = Cell_Barcode,
    values_from = count,
    values_fill = 0
  ) %>%
  # Convert Cell_Barcode to row names
  column_to_rownames("Lineage_Barcode")



#Subset enriched cells:
barcode_data_enriched <- barcode_data[barcode_data$Sample == "enriched_step1_unmapped.txt",]

#Do this for 
lineage_counts_enriched <- barcode_data_enriched %>%
  # Group by Cell_Barcode and Lineage_Barcode
  group_by(Lineage_Barcode, Cell_Barcode) %>%
  # Count unique UMIs
  summarize(count = n_distinct(UMI)) %>%
  # Ungroup
  ungroup() %>%
  # Pivot wider to create matrix-like format needed for Seurat
  pivot_wider(
    names_from = Cell_Barcode,
    values_from = count,
    values_fill = 0
  ) %>%
  # Convert Cell_Barcode to row names
  column_to_rownames("Lineage_Barcode")

#Remove original barcode data to save memory:
rm(barcode_data)

```

#Add a row of zeros for cell barcodes without any lineage barcodes, necessary to add it to the Seurat object:
```{r, include=FALSE}

#Do this for untransduced:

# Add any missing cell barcodes as columns with zeros
missing_cols <- setdiff(colnames(seurat_object_untransduced_standard_filter), colnames(lineage_counts_untransduced))
if(length(missing_cols) > 0) {
  zero_matrix <- matrix(0, 
                       nrow = nrow(lineage_counts_untransduced), 
                       ncol = length(missing_cols),
                       dimnames = list(rownames(lineage_counts_untransduced), missing_cols))
  barcode_data_untransduced_complete <- cbind(lineage_counts_untransduced, zero_matrix)
}

# Sort columns to match Seurat object order
barcode_data_untransduced_complete <- barcode_data_untransduced_complete[, colnames(seurat_object_untransduced_standard_filter)]

#Do this for enriched:

# Add any missing cell barcodes as columns with zeros
missing_cols <- setdiff(colnames(seurat_object_enriched_standard_filter), colnames(lineage_counts_enriched))
if(length(missing_cols) > 0) {
  zero_matrix <- matrix(0, 
                       nrow = nrow(lineage_counts_enriched), 
                       ncol = length(missing_cols),
                       dimnames = list(rownames(lineage_counts_enriched), missing_cols))
  barcode_data_enriched_complete <- cbind(lineage_counts_enriched, zero_matrix)
}

# Sort columns to match Seurat object order
barcode_data_enriched_complete <- barcode_data_enriched_complete[, colnames(seurat_object_enriched_standard_filter)]

```

#Add lineage barcode assay (counts) to the Seurat objects:
```{r, include=FALSE}

# Add lineage dataframe as an assay:

seurat_object_untransduced_standard_filter[["Lineage_Assay"]] <- CreateAssayObject(counts = barcode_data_untransduced_complete)

rm(barcode_data_untransduced_complete)


seurat_object_enriched_standard_filter[["Lineage_Assay"]] <- CreateAssayObject(counts = barcode_data_enriched_complete)

rm(barcode_data_enriched_complete, barcode_data_enriched, barcode_data_untransduced, lineage_counts_enriched, lineage_counts_untransduced, zero_matrix, missing_cols)

```


#Combine Seurat objects:
```{r, include=FALSE}

seurat_object_merged <- merge(seurat_object_untransduced_standard_filter, y = seurat_object_enriched_standard_filter, add.cell.ids = c("Untransduced", "Enriched"), project = "RV143")

```

#Remove original seurat objects for space:
```{r, include=FALSE}

rm(seurat_object_untransduced_standard_filter,seurat_object_enriched_standard_filter)

```

#Get metrics for barcode data before continuing with normalization:
```{r}

#Get number of unique lineage barcodes:
number_of_unique_lineage_barcodes_per_cell <- colSums(seurat_object_merged[["Lineage_Assay"]]@counts > 0)
histogram_number_of_unique_lineage_barcodes_per_cell <- table(number_of_unique_lineage_barcodes_per_cell)
plot(histogram_number_of_unique_lineage_barcodes_per_cell, xlab = "Unique Non-Zero Lineage Barcodes", ylab = "Number of Cells")

histogram_number_of_unique_lineage_barcodes_per_cell_percentage <- histogram_number_of_unique_lineage_barcodes_per_cell*100/sum(histogram_number_of_unique_lineage_barcodes_per_cell)
plot(histogram_number_of_unique_lineage_barcodes_per_cell_percentage, xlab = "Unique Non-Zero Lineage Barcodes", ylab = "% of Cells")

non_zero_lineage_barcodes <- rownames(seurat_object_merged[["Lineage_Assay"]]@counts)[rowSums(seurat_object_merged[["Lineage_Assay"]]@counts > 0) > 0]

#Make histogram of number of total umis for lineage barcodes:
plot(table(colSums(seurat_object_merged[["Lineage_Assay"]]@counts)), xlab = "Total Lineage Barcode UMIs", ylab = "Number of Cells")

#Make histogram of number of max umis for lineage barcodes:
plot(table(apply(seurat_object_merged[["Lineage_Assay"]]@counts, 2, max)), xlab = "Max Lineage Barcode UMIs", ylab = "Number of Cells")

#Get number of unique lineage barcodes for untransduced:
cells_untransduced <- colnames(seurat_object_merged)[seurat_object_merged$orig.ident == "Untransduced"]
table(colSums(seurat_object_merged[["Lineage_Assay"]]@counts[, cells_untransduced] > 0))

#Get number of unique lineage barcodes for enriched:
cells_enriched <- colnames(seurat_object_merged)[seurat_object_merged$orig.ident == "Enriched"]
table(colSums(seurat_object_merged[["Lineage_Assay"]]@counts[, cells_enriched] > 0))

print(paste("% of cells with lineage barcode with >= 1 umis:", (1-(table(colSums(seurat_object_merged[["Lineage_Assay"]]@counts))[1]/sum(table(colSums(seurat_object_merged[["Lineage_Assay"]]@counts)))))*100))

rm(number_of_unique_lineage_barcodes_per_cell, histogram_number_of_unique_lineage_barcodes_per_cell_percentage, non_zero_lineage_barcodes, cells_untransduced, cells_enriched, histogram_number_of_unique_lineage_barcodes_per_cell)

```

#Remove any cells with >= lineage_barcode_umi_threshold umis of multiple unique barcodes:
```{r, include=FALSE}

#print number of cells to check:
print(ncol(seurat_object_merged))

#17,172 cells

lineage_barcode_data <- GetAssayData(seurat_object_merged, assay = "Lineage_Assay", layer = "counts")

#Count number of lineage barcodes passing threshold:
lineage_barcodes_above_threshold <- colSums(lineage_barcode_data >= lineage_barcode_umi_threshold)

cells_without_too_many_lineage_barcodes <- names(lineage_barcodes_above_threshold)[lineage_barcodes_above_threshold < 2]

seurat_remove_cells_with_multiple_lineage_barcodes <- subset(seurat_object_merged, cells = cells_without_too_many_lineage_barcodes)

rm(seurat_object_merged, lineage_barcode_data, lineage_barcodes_above_threshold, cells_without_too_many_lineage_barcodes)

#Print current number of cells to check:
print(ncol(seurat_remove_cells_with_multiple_lineage_barcodes))

#15,975 cells

```

#Remove any cells with less than lineage_barcode_umi_threshold:
```{r, include=FALSE}

#15,975 cells

lineage_barcode_data <- GetAssayData(seurat_remove_cells_with_multiple_lineage_barcodes, assay = "Lineage_Assay", layer = "counts")

cells_with_threshold_or_more_umis_lineage_barcodes <- colnames(lineage_barcode_data)[Matrix::colSums(lineage_barcode_data >= lineage_barcode_umi_threshold) >= 1]

seurat_with_threshold_or_more_umis_lineage_barcodes <- subset(seurat_remove_cells_with_multiple_lineage_barcodes, cells = cells_with_threshold_or_more_umis_lineage_barcodes)

rm(seurat_remove_cells_with_multiple_lineage_barcodes, lineage_barcode_data, cells_with_threshold_or_more_umis_lineage_barcodes)

#Print current number of cells to check:
print(ncol(seurat_with_threshold_or_more_umis_lineage_barcodes))

#13,544 cells

```


#Remove any cells with similar number of umis of lineage barcodes (ie if 2 umis in one barcode but 3 in the other remove):
```{r, include=FALSE}

#13,544 cells

lineage_barcode_data <- GetAssayData(seurat_with_threshold_or_more_umis_lineage_barcodes, assay = "Lineage_Assay", layer = "counts")

max_second_max_diff_umi_barcodes <- apply(lineage_barcode_data, 2, function(cell) {
  sorted_umis <- sort(cell, decreasing = TRUE)
  if (length(sorted_umis) >= 2) {
    return(sorted_umis[1] - sorted_umis[2])
  } else if (length(sorted_umis) == 1) {
    return(sorted_umis[1])  # Only one barcode present
  } else {
    return(0)  # No barcodes present
  }
})

#Get examples of cells with different differences--------------------------------------------------------------------------

# Find multiple example cells for each difference value
num_examples <- 5  # Change this to show more or fewer examples

cells_with_diff_0 <- names(max_second_max_diff_umi_barcodes)[max_second_max_diff_umi_barcodes == 0][1:num_examples]
cells_with_diff_1 <- names(max_second_max_diff_umi_barcodes)[max_second_max_diff_umi_barcodes == 1][1:num_examples]
cells_with_diff_2 <- names(max_second_max_diff_umi_barcodes)[max_second_max_diff_umi_barcodes == 2][1:num_examples]
cells_with_diff_3 <- names(max_second_max_diff_umi_barcodes)[max_second_max_diff_umi_barcodes == 3][1:num_examples]

# Print examples for each category
cat("========================================\n")
cat("=== Cells with UMI difference = 0 ===\n")
cat("========================================\n")
for (cell in cells_with_diff_0) {
  print_non_zero_barcodes(cell, lineage_barcode_data, 0)
}

cat("========================================\n")
cat("=== Cells with UMI difference = 1 ===\n")
cat("========================================\n")
for (cell in cells_with_diff_1) {
  print_non_zero_barcodes(cell, lineage_barcode_data, 1)
}

cat("========================================\n")
cat("=== Cells with UMI difference = 2 ===\n")
cat("========================================\n")
for (cell in cells_with_diff_2) {
  print_non_zero_barcodes(cell, lineage_barcode_data, 2)
}


cat("========================================\n")
cat("=== Cells with UMI difference = 3 ===\n")
cat("========================================\n")
for (cell in cells_with_diff_3) {
  print_non_zero_barcodes(cell, lineage_barcode_data, 3)
}

# Look at cells with highest umis by sorting cells by highest max UMI within each difference category
max_umis <- apply(lineage_barcode_data, 2, function(cell) max(cell))

cells_with_diff_0 <- names(sort(max_umis[cells_with_diff_0[!is.na(cells_with_diff_0)]], decreasing = TRUE))
cells_with_diff_1 <- names(sort(max_umis[cells_with_diff_1[!is.na(cells_with_diff_1)]], decreasing = TRUE))
cells_with_diff_2 <- names(sort(max_umis[cells_with_diff_2[!is.na(cells_with_diff_2)]], decreasing = TRUE))
cells_with_diff_3 <- names(sort(max_umis[cells_with_diff_3[!is.na(cells_with_diff_3)]], decreasing = TRUE))

# Print examples for each category
cat("========================================\n")
cat("=== Cells with UMI difference = 0 ===\n")
cat("========================================\n")
for (cell in cells_with_diff_0) {
  print_non_zero_barcodes(cell, lineage_barcode_data, 0)
}
cat("========================================\n")
cat("=== Cells with UMI difference = 1 ===\n")
cat("========================================\n")
for (cell in cells_with_diff_1) {
  print_non_zero_barcodes(cell, lineage_barcode_data, 1)
}
cat("========================================\n")
cat("=== Cells with UMI difference = 2 ===\n")
cat("========================================\n")
for (cell in cells_with_diff_2) {
  print_non_zero_barcodes(cell, lineage_barcode_data, 2)
}
cat("========================================\n")
cat("=== Cells with UMI difference = 3 ===\n")
cat("========================================\n")
for (cell in cells_with_diff_3) {
  print_non_zero_barcodes(cell, lineage_barcode_data, 3)
}

print_non_zero_barcodes("Untransduced_AAACCCGCATCCGCAT-1", lineage_barcode_data)

#-------------------------------------------------------------------------------------

#Remove cells within a distance of lineage_barcode_umi_threshold-1 between top lineage barcode umis:

#Get cells with greater than lineage_barcode_umi_threshold-1 lineage barcode umis (aka lineage_barcode_umi_threshold or greater distance):
cells_with_high_difference_between_barcodes <- names(max_second_max_diff_umi_barcodes)[max_second_max_diff_umi_barcodes >= lineage_barcode_umi_threshold]

seurat_strict_barcode_requirements <- subset(seurat_with_threshold_or_more_umis_lineage_barcodes, cells = cells_with_high_difference_between_barcodes)

rm(seurat_with_threshold_or_more_umis_lineage_barcodes, max_second_max_diff_umi_barcodes, num_examples, cells_with_diff_0, cells_with_diff_1, cells_with_diff_2, cells_with_diff_3, max_umis, lineage_barcode_data, cell, cells_with_high_difference_between_barcodes)

#Print current number of cells to check:
print(ncol(seurat_strict_barcode_requirements))

#12,996 cells

```

#Assign lineage barcode with the most umis to each cell:
```{r, include=FALSE}

lineage_barcode_data <- GetAssayData(seurat_strict_barcode_requirements, assay = "Lineage_Assay", layer = "counts")

seurat_strict_barcode_requirements$Lineage_Barcode_w_Max_Count <- apply(lineage_barcode_data, 2, function(x) {
  rownames(lineage_barcode_data)[which.max(x)]
})

rm(lineage_barcode_data)

```

#Load enrichment mini-library for assignment of lineage barcodes:
```{r, include=FALSE}

#Get gibson sequences:
gibson_assembly_sequences <- read.table("2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Organized_Data/Gibson_Assembly_Sequences_20241212.csv")

#Isolate just the barcodes:
mini_library_barcodes <- apply(gibson_assembly_sequences, 1, function(x) {
  substring(x, first = 50, last = 69)
})

```

#Assign enrichment mini-library positive or not to cells in the seurat object:
```{r, include=FALSE}

seurat_strict_barcode_requirements$has_targeted_lineage_barcode <- ifelse((seurat_strict_barcode_requirements$Lineage_Barcode_w_Max_Count %in% mini_library_barcodes),
                              "Targeted_Lineage_Barcode",                                           
                              "Not_Targeted_Lineage_Barcode")

#Create a separate metadata column for just enrichment mini-library assigned barcodes:
seurat_strict_barcode_requirements$Mini_Library_Assignment <-
  ifelse(seurat_strict_barcode_requirements$has_targeted_lineage_barcode == "Targeted_Lineage_Barcode",
         seurat_strict_barcode_requirements$Lineage_Barcode_w_Max_Count,
         "Non-Enriched")

```


#Normalize data with strict barcoding requirements:
```{r, include=FALSE}

seurat_strict_barcode_requirements <- JoinLayers(seurat_strict_barcode_requirements)

seurat_strict_barcode_requirements <- NormalizeData(seurat_strict_barcode_requirements,
              normalization.method = "LogNormalize",
              scale.factor = 10000)

seurat_strict_barcode_requirements <- ScaleData(seurat_strict_barcode_requirements)

seurat_strict_barcode_requirements <- FindVariableFeatures(seurat_strict_barcode_requirements)

```

#Run dimension reductions:
```{r}

seurat_strict_barcode_requirements <- RunPCA(seurat_strict_barcode_requirements, features = VariableFeatures(object = seurat_strict_barcode_requirements))

seurat_strict_barcode_requirements <- FindNeighbors(seurat_strict_barcode_requirements, dims = 1:10)
seurat_strict_barcode_requirements <- FindClusters(seurat_strict_barcode_requirements, resolution = 0.5)

seurat_strict_barcode_requirements <- RunUMAP(seurat_strict_barcode_requirements, dims = 1:10)

DimPlot(seurat_strict_barcode_requirements, reduction = "umap")

DimPlot(seurat_strict_barcode_requirements, reduction = "umap", group.by = "Mini_Library_Assignment", ) + NoLegend()

```

#Save strict barcode requirement seurat object:
```{r, include=FALSE}

getwd()

saveRDS(seurat_strict_barcode_requirements,"RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Create_Seurat_Objects/seurat_strict_barcode_requirements")

```


#Save raw reads for cNMF:
```{r, include=FALSE}

getwd()

seurat_object_strict_barcode_requirements_raw_count <- seurat_strict_barcode_requirements[["RNA"]]$counts

writeMM(seurat_object_strict_barcode_requirements_raw_count,"RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Create_Seurat_Objects/seurat_object_strict_barcode_requirements_raw_count/matrix.mtx")

# Output cell barcodes
barcodes <- colnames(seurat_object_strict_barcode_requirements_raw_count)
write.table(as.data.frame(barcodes), "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Create_Seurat_Objects/seurat_object_strict_barcode_requirements_raw_count/barcodes.tsv",
            col.names = FALSE, row.names = FALSE, sep = "\t")


# Output feature names
gene_names <- rownames(seurat_object_strict_barcode_requirements_raw_count)
features <- data.frame("gene_id" = gene_names,"gene_name" = gene_names,type = "Gene Expression")
write.table(as.data.frame(features), sep = "\t", "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Create_Seurat_Objects/seurat_object_strict_barcode_requirements_raw_count/features.tsv",
            col.names = FALSE, row.names = FALSE)

#Gzip files for cNMF:
system("gzip -f RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Create_Seurat_Objects/seurat_object_strict_barcode_requirements_raw_count/matrix.mtx")
system("gzip -f RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Create_Seurat_Objects/seurat_object_strict_barcode_requirements_raw_count/barcodes.tsv")
system("gzip -f RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Create_Seurat_Objects/seurat_object_strict_barcode_requirements_raw_count/features.tsv")

```




