---
title: "Enriched_Lineage_Comparison_Analysis"
output: html_document
date: "2026-02-05"
---

```{r setup, include=FALSE}

#Run only in chunk:
knitr::opts_knit$set(root.dir = '/Users/robertvandervelde/Library/CloudStorage/GoogleDrive-robertvandervelde@sydshafferlab.com/My Drive/Vander_Velde_Share/Experiments/RV143')

getwd()

```


#Load Libraries:
```{r, include=FALSE}

rm(list = ls())
library(Seurat)
library(tidyverse)
library(ggplot2)
library(tidyplots)
library(dplyr)

getwd()

set.seed(1)

```

#Set variable:
```{r, include=FALSE}

font_size <- 8
font_type <- "Helvetica Neue"

panel_dimensions <- c(30,30)

```

#Load seurat object:
```{r, include=FALSE}

seurat_strict_barcode_requirements <- readRDS("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Create_Seurat_Objects/seurat_strict_barcode_requirements")

```

#Get only cells in enrichment sample:
```{r, include=FALSE}

Idents(seurat_strict_barcode_requirements) <- "orig.ident"

seurat_strict_barcode_requirements_enrichment_only <- subset(x = seurat_strict_barcode_requirements, idents = "Enriched")
  
rm(seurat_strict_barcode_requirements)

```



#Isolate lineage with 3 or more cells:
```{r, include=FALSE}

#Get list of lineage with 3 or more cells:
lineage_list <- names(table(seurat_strict_barcode_requirements_enrichment_only$Mini_Library_Assignment)[table(seurat_strict_barcode_requirements_enrichment_only$Mini_Library_Assignment) >= 3])

#Remove "Non-Enriched" from lineage list:
lineage_list <- lineage_list[lineage_list != "Non-Enriched"]

Idents(seurat_strict_barcode_requirements_enrichment_only) <- "Mini_Library_Assignment"

#Subset based on list:
seurat_large_lineages <- subset(x = seurat_strict_barcode_requirements_enrichment_only, idents = lineage_list)

#Remove large seurat object:
rm(seurat_strict_barcode_requirements_enrichment_only)

#Reset idents:
Idents(seurat_large_lineages) <- "seurat_clusters"

```

#Run FindAllMarkers comparing each lineage to all others:
```{r, include=FALSE}

Idents(seurat_large_lineages) <- "Mini_Library_Assignment"

interlineage_findmarkers <- FindAllMarkers(seurat_large_lineages)

```

#Save FindAllMarkers table:
```{r, include=FALSE}

write.csv2(interlineage_findmarkers, "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Enriched_Lineage_Comparison_Analysis/FindAllMarkers.csv")

```

#Create label list, with the number of cells in each lineage:
```{r, include=FALSE}

dictionary_labels <- table(seurat_large_lineages$Mini_Library_Assignment) %>% 
  {setNames(paste0("Lineage ", seq_along(.), " (cells = ", ., ")"), names(.))}

```

#Create dataframe for IL1RAPL1 expression:

```{r, include=FALSE}

IL1RAPL1_df <- FetchData(seurat_large_lineages, 
                      vars = c("IL1RAPL1", "Mini_Library_Assignment"),
                      layer = 'data')

```

#Plot IL1RAPL1:
```{r}

p <- tidyplot(data = IL1RAPL1_df, x = Mini_Library_Assignment, y = IL1RAPL1) |>
  add_data_points_beeswarm(shape = 16, size = 0.5) |>
  add_boxplot(show_outliers = FALSE) |>
  adjust_size(width = panel_dimensions[1], height = panel_dimensions[2]) |>
  adjust_x_axis(rotate_labels = 90,
                labels = dictionary_labels) |>
  adjust_x_axis_title("") |>
  adjust_y_axis_title("Normalized IL1RAPL1\nExpression") |>
  adjust_y_axis(limits = c(NA, max(IL1RAPL1_df$IL1RAPL1, na.rm = TRUE) * 1.15)) |>
  adjust_font(fontsize = font_size, family = font_type) +
  ggplot2::theme(panel.border = element_rect(color = "black", linewidth = 0.5),
                 axis.line = element_blank())

ggsave(p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Enriched_Lineage_Comparison/IL1RAPL1.pdf", device = cairo_pdf)
    
  
```


#Create dataframe for NRG1 expression:

```{r, include=FALSE}

NRG1_df <- FetchData(seurat_large_lineages, 
                      vars = c("NRG1", "Mini_Library_Assignment"),
                      layer = 'data')

```

#Plot NRG1:
```{r}

p <- tidyplot(data = NRG1_df, x = Mini_Library_Assignment, y = NRG1) |>
  add_data_points_beeswarm(shape = 16, size = 0.5) |>
  add_boxplot(show_outliers = FALSE) |>
  adjust_size(width = panel_dimensions[1], height = panel_dimensions[2]) |>
  adjust_x_axis(rotate_labels = 90,
                labels = dictionary_labels) |>
  adjust_x_axis_title("") |>
  adjust_y_axis_title("Normalized NRG1\nExpression") |>
  adjust_y_axis(limits = c(NA, max(NRG1_df$NRG1, na.rm = TRUE) * 1.15)) |>
  adjust_font(fontsize = font_size, family = font_type) +
  ggplot2::theme(panel.border = element_rect(color = "black", linewidth = 0.5),
                 axis.line = element_blank())

ggsave(p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Enriched_Lineage_Comparison/NRG1.pdf", device = cairo_pdf)
    
  
```


#Create dataframe for HS6ST2 expression:

```{r, include=FALSE}

HS6ST2_df <- FetchData(seurat_large_lineages, 
                      vars = c("HS6ST2", "Mini_Library_Assignment"),
                      layer = 'data')

```

#Plot HS6ST2:
```{r}

p <- tidyplot(data = HS6ST2_df, x = Mini_Library_Assignment, y = HS6ST2) |>
  add_data_points_beeswarm(shape = 16, size = 0.5) |>
  add_boxplot(show_outliers = FALSE) |>
  adjust_size(width = panel_dimensions[1], height = panel_dimensions[2]) |>
  adjust_x_axis(rotate_labels = 90,
                labels = dictionary_labels) |>
  adjust_x_axis_title("") |>
  adjust_y_axis_title("Normalized HS6ST2\nExpression") |>
  adjust_y_axis(limits = c(NA, max(HS6ST2_df$HS6ST2, na.rm = TRUE) * 1.15)) |>
  adjust_font(fontsize = font_size, family = font_type) +
  ggplot2::theme(panel.border = element_rect(color = "black", linewidth = 0.5),
                 axis.line = element_blank())

ggsave(p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Enriched_Lineage_Comparison/HS6ST2.pdf", device = cairo_pdf)
    
  
```

