---
title: "Plot_Gene_Expression"
output: html_document
date: "2025-06-15"
---


```{r setup, include=FALSE}

#Run only in chunk:
knitr::opts_knit$set(root.dir = '/Users/robertvandervelde/Library/CloudStorage/GoogleDrive-robertvandervelde@sydshafferlab.com/My Drive/Vander_Velde_Share/Experiments/RV143')

getwd()

```


#Load Libraries:
```{r, include=FALSE}

rm(list = ls())
library(Seurat)
library(tidyverse)
library(Matrix)

library(egg)

library(Cairo)

set.seed(1)

```


#Set font size and type:
```{r, include=FALSE}

font_size <- 8
font_type <- "Helvetica Neue"
point_size <- 0.01
jitter_width <- 0.2
point_stroke_weight <- 0.05
box_line_width <- 0.2
nudge_size <- 0.25

panel_dimensions <- c(0.4,1)

```

#Load strict barcode requirements seurat object:
```{r, include=FALSE}

seurat_strict_barcode_requirements <- readRDS("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Create_Seurat_Objects/seurat_strict_barcode_requirements")


```

#Load FindMarkers to check significance:
```{r, include=FALSE}


markers_untranduced_vs_transduced <- read.csv("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Compare_Differential_Expression/markers_untranduced_vs_transduced.csv", row.names = 1)

markers_targeted_vs_not <- read.csv("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Compare_Differential_Expression/markers_targeted_vs_not.csv", row.names = 1)

```

#Subset to only look at enriched set:
```{r, include=FALSE}

enriched_only <- subset(x=seurat_strict_barcode_requirements, subset = orig.ident == "Enriched")

```

#Find markers for all cells assigned a lineage barcode:
```{r, include=FALSE}

#Change idents to be linage barcode assignment:
Idents(enriched_only) <- enriched_only$has_targeted_lineage_barcode

```

#Isolate data for DCT:
```{r}

#See if DCT is significant:
print(markers_targeted_vs_not["DCT",])

#Create dataframe for DCT expression:
DCT_df <- FetchData(enriched_only, 
                      vars = c("DCT", "Mini_Library_Assignment","has_targeted_lineage_barcode"),
                      layer = 'data')

#Save row names:
original_rownames <- rownames(DCT_df)

#Identify outliers for plotting:
DCT_df <- DCT_df %>%
  group_by(has_targeted_lineage_barcode) %>%
  mutate(is_outlier = DCT < quantile(DCT, 0.25) - 1.5*IQR(DCT) | 
                      DCT > quantile(DCT, 0.75) + 1.5*IQR(DCT)) %>%
  ungroup() %>%
  as.data.frame()

# Restore row names
rownames(DCT_df) <- original_rownames

#Remove row names list:
rm(original_rownames)

```

#Plot DCT as a simple boxplot:
```{r}

p <- ggplot(DCT_df,
       aes(x = factor(has_targeted_lineage_barcode, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = DCT,
           color = factor(has_targeted_lineage_barcode))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(DCT_df, is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("Normalized Expression") +
  ggtitle(paste("DCT\nAdj. p-value:\n", format(markers_targeted_vs_not["DCT",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_Gene_Expression/DCT_Box_Plot_Only.pdf", device = cairo_pdf, width = 2, height = 3)
  
```

#Remove DCT dataframe:
```{r, include=FALSE}

rm(DCT_df)

```



#Isolate data for PMEL:
```{r}

#See if PMEL is significant:
print(markers_targeted_vs_not["PMEL",])

#Create dataframe for PMEL expression:
PMEL_df <- FetchData(enriched_only, 
                      vars = c("PMEL", "Mini_Library_Assignment","has_targeted_lineage_barcode"),
                      layer = 'data')

#Save row names:
original_rownames <- rownames(PMEL_df)

#Identify outliers for plotting:
PMEL_df <- PMEL_df %>%
  group_by(has_targeted_lineage_barcode) %>%
  mutate(is_outlier = PMEL < quantile(PMEL, 0.25) - 1.5*IQR(PMEL) | 
                      PMEL > quantile(PMEL, 0.75) + 1.5*IQR(PMEL)) %>%
  ungroup() %>%
  as.data.frame()

# Restore row names
rownames(PMEL_df) <- original_rownames

#Remove row names list:
rm(original_rownames)

```

#Plot PMEL as a simple boxplot:
```{r}

p <- ggplot(PMEL_df,
       aes(x = factor(has_targeted_lineage_barcode, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = PMEL,
           color = factor(has_targeted_lineage_barcode))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(PMEL_df, is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("Normalized Expression") +
  ggtitle(paste("PMEL\nAdj. p-value:\n", format(markers_targeted_vs_not["PMEL",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_Gene_Expression/PMEL_Box_Plot_Only.pdf", device = cairo_pdf, width = 2, height = 3)
  
```

#Remove PMEL dataframe:
```{r, include=FALSE}

rm(PMEL_df)

```



#Isolate data for IGFBP7:
```{r}


#See if IGFBP7 is significant:
print(markers_targeted_vs_not["IGFBP7",])

#Create dataframe for IGFBP7 expression:
IGFBP7_df <- FetchData(enriched_only, 
                      vars = c("IGFBP7", "Mini_Library_Assignment","has_targeted_lineage_barcode"),
                      layer = 'data')

#Save row names:
original_rownames <- rownames(IGFBP7_df)

#Identify outliers for plotting:
IGFBP7_df <- IGFBP7_df %>%
  group_by(has_targeted_lineage_barcode) %>%
  mutate(is_outlier = IGFBP7 < quantile(IGFBP7, 0.25) - 1.5*IQR(IGFBP7) | 
                      IGFBP7 > quantile(IGFBP7, 0.75) + 1.5*IQR(IGFBP7)) %>%
  ungroup() %>%
  as.data.frame()

# Restore row names
rownames(IGFBP7_df) <- original_rownames

#Remove row names list:
rm(original_rownames)

```

#Plot IGFBP7 as a simple boxplot:
```{r}

p <- ggplot(IGFBP7_df,
       aes(x = factor(has_targeted_lineage_barcode, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = IGFBP7,
           color = factor(has_targeted_lineage_barcode))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(IGFBP7_df, is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("Normalized Expression") +
  ggtitle("IGFBP7") +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_Gene_Expression/IGFBP7_Box_Plot_Only.pdf", device = cairo_pdf, width = 2, height = 3)
  
```

#Remove IGFBP7 dataframe:
```{r, include=FALSE}

rm(IGFBP7_df)

```






#Isolate data for TYR:
```{r}

#See if TYR is significant:
print(markers_targeted_vs_not["TYR",])

#Create dataframe for TYR expression:
TYR_df <- FetchData(enriched_only, 
                      vars = c("TYR", "Mini_Library_Assignment","has_targeted_lineage_barcode"),
                      layer = 'data')

#Save row names:
original_rownames <- rownames(TYR_df)

#Identify outliers for plotting:
TYR_df <- TYR_df %>%
  group_by(has_targeted_lineage_barcode) %>%
  mutate(is_outlier = TYR < quantile(TYR, 0.25) - 1.5*IQR(TYR) | 
                      TYR > quantile(TYR, 0.75) + 1.5*IQR(TYR)) %>%
  ungroup() %>%
  as.data.frame()

# Restore row names
rownames(TYR_df) <- original_rownames

#Remove row names list:
rm(original_rownames)

```

#Plot TYR as a simple boxplot:
```{r}

p <- ggplot(TYR_df,
       aes(x = factor(has_targeted_lineage_barcode, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = TYR,
           color = factor(has_targeted_lineage_barcode))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(TYR_df, is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("Normalized Expression") +
  ggtitle(paste("TYR\nAdj. p-value:\n", format(markers_targeted_vs_not["TYR",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_Gene_Expression/TYR_Box_Plot_Only.pdf", device = cairo_pdf, width = 2, height = 3)
  
```

#Remove TYR dataframe:
```{r, include=FALSE}

rm(TYR_df)

```




#Isolate data for CD44:
```{r}

#See if CD44 is significant:
print(markers_targeted_vs_not["CD44",])

#Create dataframe for CD44 expression:
CD44_df <- FetchData(enriched_only, 
                      vars = c("CD44", "Mini_Library_Assignment","has_targeted_lineage_barcode"),
                      layer = 'data')

#Save row names:
original_rownames <- rownames(CD44_df)

#Identify outliers for plotting:
CD44_df <- CD44_df %>%
  group_by(has_targeted_lineage_barcode) %>%
  mutate(is_outlier = CD44 < quantile(CD44, 0.25) - 1.5*IQR(CD44) | 
                      CD44 > quantile(CD44, 0.75) + 1.5*IQR(CD44)) %>%
  ungroup() %>%
  as.data.frame()

# Restore row names
rownames(CD44_df) <- original_rownames

#Remove row names list:
rm(original_rownames)

```

#Plot CD44 as a simple boxplot:
```{r}

p <- ggplot(CD44_df,
       aes(x = factor(has_targeted_lineage_barcode, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = CD44,
           color = factor(has_targeted_lineage_barcode))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(CD44_df, is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("Normalized Expression") +
  ggtitle(paste("CD44\nAdj. p-value:\n", format(markers_targeted_vs_not["CD44",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_Gene_Expression/CD44_Box_Plot_Only.pdf", device = cairo_pdf, width = 2, height = 3)
  
```


#Remove CD44 dataframe:
```{r, include=FALSE}

rm(CD44_df)

```






#Isolate data for CD59:
```{r}

#See if CD59 is significant:
print(markers_targeted_vs_not["CD59",])

#Create dataframe for CD59 expression:
CD59_df <- FetchData(enriched_only, 
                      vars = c("CD59", "Mini_Library_Assignment","has_targeted_lineage_barcode"),
                      layer = 'data')

#Save row names:
original_rownames <- rownames(CD59_df)

#Identify outliers for plotting:
CD59_df <- CD59_df %>%
  group_by(has_targeted_lineage_barcode) %>%
  mutate(is_outlier = CD59 < quantile(CD59, 0.25) - 1.5*IQR(CD59) | 
                      CD59 > quantile(CD59, 0.75) + 1.5*IQR(CD59)) %>%
  ungroup() %>%
  as.data.frame()

# Restore row names
rownames(CD59_df) <- original_rownames

#Remove row names list:
rm(original_rownames)

```

#Plot CD59 as a simple boxplot:
```{r}

p <- ggplot(CD59_df,
       aes(x = factor(has_targeted_lineage_barcode, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = CD59,
           color = factor(has_targeted_lineage_barcode))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(CD59_df, is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("Normalized Expression") +
  ggtitle(paste("CD59\nAdj. p-value:\n", format(markers_targeted_vs_not["CD59",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_Gene_Expression/CD59_Box_Plot_Only.pdf", device = cairo_pdf, width = 2, height = 3)
  
```


#Remove CD59 dataframe:
```{r, include=FALSE}

rm(CD59_df)

```




#Isolate data for HRAS:
```{r}

#See if HRAS is significant:
print(markers_targeted_vs_not["HRAS",])

#Create dataframe for HRAS expression:
HRAS_df <- FetchData(enriched_only, 
                      vars = c("HRAS", "Mini_Library_Assignment","has_targeted_lineage_barcode"),
                      layer = 'data')

#Save row names:
original_rownames <- rownames(HRAS_df)

#Identify outliers for plotting:
HRAS_df <- HRAS_df %>%
  group_by(has_targeted_lineage_barcode) %>%
  mutate(is_outlier = HRAS < quantile(HRAS, 0.25) - 1.5*IQR(HRAS) | 
                      HRAS > quantile(HRAS, 0.75) + 1.5*IQR(HRAS)) %>%
  ungroup() %>%
  as.data.frame()

# Restore row names
rownames(HRAS_df) <- original_rownames

#Remove row names list:
rm(original_rownames)

```

#Plot HRAS as a simple boxplot:
```{r}

p <- ggplot(HRAS_df,
       aes(x = factor(has_targeted_lineage_barcode, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = HRAS,
           color = factor(has_targeted_lineage_barcode))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(HRAS_df, is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("Normalized Expression") +
  ggtitle(paste("HRAS\nAdj. p-value:\n", format(markers_targeted_vs_not["HRAS",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_Gene_Expression/HRAS_Box_Plot_Only.pdf", device = cairo_pdf, width = 2, height = 3)
  
```

#Remove HRAS dataframe:
```{r, include=FALSE}

rm(HRAS_df)

```




#Isolate data for MYOF:
```{r}

#See if MYOF is significant:
print(markers_targeted_vs_not["MYOF",])

#Create dataframe for MYOF expression:
MYOF_df <- FetchData(enriched_only, 
                      vars = c("MYOF", "Mini_Library_Assignment","has_targeted_lineage_barcode"),
                      layer = 'data')

#Save row names:
original_rownames <- rownames(MYOF_df)

#Identify outliers for plotting:
MYOF_df <- MYOF_df %>%
  group_by(has_targeted_lineage_barcode) %>%
  mutate(is_outlier = MYOF < quantile(MYOF, 0.25) - 1.5*IQR(MYOF) | 
                      MYOF > quantile(MYOF, 0.75) + 1.5*IQR(MYOF)) %>%
  ungroup() %>%
  as.data.frame()

# Restore row names
rownames(MYOF_df) <- original_rownames

#Remove row names list:
rm(original_rownames)

```

#Plot MYOF as a simple boxplot:
```{r}

p <- ggplot(MYOF_df,
       aes(x = factor(has_targeted_lineage_barcode, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = MYOF,
           color = factor(has_targeted_lineage_barcode))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(MYOF_df, is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("Normalized Expression") +
  ggtitle(paste("MYOF\nAdj. p-value:\n", format(markers_targeted_vs_not["MYOF",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_Gene_Expression/MYOF_Box_Plot_Only.pdf", device = cairo_pdf, width = 2, height = 3)
  
```

#Remove MYOF dataframe:
```{r, include=FALSE}

rm(MYOF_df)

```




#Isolate data for MKI67:
```{r}

#See if MKI67 is significant:
print(markers_targeted_vs_not["MKI67",])

#Create dataframe for MKI67 expression:
MKI67_df <- FetchData(enriched_only, 
                      vars = c("MKI67", "Mini_Library_Assignment","has_targeted_lineage_barcode"),
                      layer = 'data')

#Save row names:
original_rownames <- rownames(MKI67_df)

#Identify outliers for plotting:
MKI67_df <- MKI67_df %>%
  group_by(has_targeted_lineage_barcode) %>%
  mutate(is_outlier = MKI67 < quantile(MKI67, 0.25) - 1.5*IQR(MKI67) | 
                      MKI67 > quantile(MKI67, 0.75) + 1.5*IQR(MKI67)) %>%
  ungroup() %>%
  as.data.frame()

# Restore row names
rownames(MKI67_df) <- original_rownames

#Remove row names list:
rm(original_rownames)

```

#Plot MKI67 as a simple boxplot:
```{r}

p <- ggplot(MKI67_df,
       aes(x = factor(has_targeted_lineage_barcode, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = MKI67,
           color = factor(has_targeted_lineage_barcode))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(MKI67_df, is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("Normalized Expression") +
  ggtitle(paste("MKI67\nAdj. p-value:\n", format(markers_targeted_vs_not["MKI67",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_Gene_Expression/MKI67_Box_Plot_Only.pdf", device = cairo_pdf, width = 2, height = 3)
  
```


#Remove MKI67 dataframe:
```{r, include=FALSE}

rm(MKI67_df)

```





#Isolate data for MLANA:
```{r}

#See if MLANA is significant:
print(markers_targeted_vs_not["MLANA",])

#Create dataframe for MLANA expression:
MLANA_df <- FetchData(enriched_only, 
                      vars = c("MLANA", "Mini_Library_Assignment","has_targeted_lineage_barcode"),
                      layer = 'data')

#Save row names:
original_rownames <- rownames(MLANA_df)

#Identify outliers for plotting:
MLANA_df <- MLANA_df %>%
  group_by(has_targeted_lineage_barcode) %>%
  mutate(is_outlier = MLANA < quantile(MLANA, 0.25) - 1.5*IQR(MLANA) | 
                      MLANA > quantile(MLANA, 0.75) + 1.5*IQR(MLANA)) %>%
  ungroup() %>%
  as.data.frame()

# Restore row names
rownames(MLANA_df) <- original_rownames

#Remove row names list:
rm(original_rownames)

```


#Plot MLANA as a simple boxplot:
```{r}

p <- ggplot(MLANA_df,
       aes(x = factor(has_targeted_lineage_barcode, 
                     levels = c("Not_Targeted_Lineage_Barcode", "Targeted_Lineage_Barcode"),
                     labels = c("Non-Enriched", "Enriched")), 
           y = MLANA,
           color = factor(has_targeted_lineage_barcode))) +
  geom_boxplot( outlier.shape = NA,
                width = jitter_width/0.533) +
  geom_jitter(data = subset(MLANA_df, is_outlier == TRUE),
              width = jitter_width,
              height = 0,
              size = point_size,
              shape = 16) + #Set outliers to dot without border
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    plot.title = element_text(family = font_type, size = font_size, color = "black", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none"
  ) +
  xlab("") +
  ylab("Normalized Expression") +
  ggtitle(paste("MLANA\nAdj. p-value:\n", format(markers_targeted_vs_not["MLANA",]$p_val_adj, scientific = TRUE, digits = 3))) +
  scale_color_manual(values = c("Not_Targeted_Lineage_Barcode" = "gray60", 
                              "Targeted_Lineage_Barcode" = "#CC6677"), guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05)))

p

#Set explicit panel size using egg package:
p <- set_panel_size(p, 
                    width  = unit(panel_dimensions[1], "inches"),
                    height = unit(panel_dimensions[2], "inches"))

ggsave(plot = p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_Gene_Expression/MLANA_Box_Plot_Only.pdf", device = cairo_pdf, width = 2, height = 3)
  
```


#Remove MLANA dataframe:
```{r, include=FALSE}

rm(MLANA_df)

```
