---
title: "Plot_Percentage_of_Cells"
output: html_document
date: "2025-07-01"
---


```{r setup, include=FALSE}

#Run only in chunk:
knitr::opts_knit$set(root.dir = '/Users/robertvandervelde/Library/CloudStorage/GoogleDrive-robertvandervelde@sydshafferlab.com/My Drive/Vander_Velde_Share/Experiments/RV143')

getwd()

```


#Load Libraries:
```{r, include=FALSE}

rm(list = ls())
library(Seurat)
library(tidyverse)
library(Matrix)
library(dplyr)
library(stringr)
library(reshape2)
library(ggforce)
library(ggbreak)

set.seed(1)

```

#Define functions:

```{r, include=FALSE}

#For reverse complementing:
revcomp <- function(seq) {
  chartr("ATCG", "TAGC", paste(rev(strsplit(seq, "")[[1]]), collapse=""))
}

```


#Set variable:
```{r, include=FALSE}

font_size <- 8
font_type <- "Helvetica Neue"

```



#Load strict barcode requirements seurat object:
```{r, include=FALSE}

seurat_strict_barcode_requirements <- readRDS("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Create_Seurat_Objects/seurat_strict_barcode_requirements")

```

#Get percentages of cells in each orig.ident (untransduce and transduced/enriched) with a specific targeted barcode:
```{r, include=FALSE}

untransduced_table <- prop.table(table(seurat_strict_barcode_requirements@meta.data[seurat_strict_barcode_requirements@meta.data$orig.ident == "Untransduced", "Mini_Library_Assignment"], useNA = "always"))*100

transduced_table <- prop.table(table(seurat_strict_barcode_requirements@meta.data[seurat_strict_barcode_requirements@meta.data$orig.ident == "Enriched", "Mini_Library_Assignment"], useNA = "always"))*100

#Remove NA columns:
untransduced_table <- untransduced_table[!is.na(names(untransduced_table))]
transduced_table <- transduced_table[!is.na(names(transduced_table))]


```

#Load ALL targeted barcodes:
```{r, include=FALSE}

#Get gibson sequences:
gibson_assembly_sequences <- read.table("2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Organized_Data/Gibson_Assembly_Sequences_20241212.csv")

#Isolate just the barcodes:
mini_library_barcodes <- apply(gibson_assembly_sequences, 1, function(x) {
  substring(x, first = 50, last = 69)
})

```

#Load the gDNA barcode data:
```{r, include=FALSE}

gDNA_counts <- read.csv("2024_10_23_gDNA_Sequencing/Step3_40bp_Scaffold_90Match_3mm/Post_Cluster_Analysis/Organized_Data/counts_without_spike_ins.csv", row.names = 1)

rownames(gDNA_counts) <- lapply(rownames(gDNA_counts),revcomp)

```

#Get proportion of gDNA reads:
```{r, include=FALSE}

gDNA_targeted_percentages <- gDNA_counts$Pre2[rownames(gDNA_counts) %in% mini_library_barcodes] * 100 / sum(gDNA_counts$Pre2)

names(gDNA_targeted_percentages) <- rownames(gDNA_counts)[rownames(gDNA_counts) %in% mini_library_barcodes]

```

#Combine percentages into a single dataframe:
```{r, include=FALSE}

#Create vectors with 0s for missing lineage barcodes:
untransduced_vector <- rep(0, length(mini_library_barcodes))
names(untransduced_vector) <- mini_library_barcodes
untransduced_vector[names(untransduced_table)] <- unlist(untransduced_table)
#Remove non-enriched:
untransduced_vector <- untransduced_vector[names(untransduced_vector) != "Non-Enriched"]

transduced_vector <- rep(0, length(mini_library_barcodes))
names(transduced_vector) <- mini_library_barcodes
transduced_vector[names(transduced_table)] <- unlist(transduced_table)
#Remove non-enriched:
transduced_vector <- transduced_vector[names(transduced_vector) != "Non-Enriched"]

#Create gDNA vector:
gDNA_vector <- rep(0, length(mini_library_barcodes))
names(gDNA_vector) <- mini_library_barcodes
gDNA_vector[names(gDNA_targeted_percentages)] <- unlist(gDNA_targeted_percentages)

all_percentages_dataframe <- data.frame(Untransduced = untransduced_vector,
                                        Transduced = transduced_vector,
                                        gDNA = gDNA_vector,
                                        row.names = mini_library_barcodes)

```


#Plot the percentages
```{r}

ggplot(all_percentages_dataframe, aes(x = Untransduced, y = Transduced, color = gDNA)) +
  geom_point(size = 2) +
  labs(x = "% of Cells in Untransduced", 
       y = "% of Cells in Transduced",
       color = "% of\nPre-Treatment\ngDNA Reads") +
  scale_color_gradient(low = "gray", high = "black") +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.text = element_text(family = font_type, size = font_size, color = "black")
  )

```

#Plot with gDNA on the x-axis, so that all points are visible:
```{r}

ggplot(all_percentages_dataframe, aes(x = gDNA, y = Transduced, color = Untransduced)) +
  geom_point(size = 0.5) +
  labs(x = "% of Pre-Treatment gDNA Reads",
       y = "% of Cells in Transduced",
       color = "% of Cells\nin Untransduced") +
  scale_color_gradient(low = "gray", high = "black") +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_text(angle = 90, vjust = 0.5),
    aspect.ratio = 1
  )


ggsave("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_Percentage_of_Cells/Untransduced_as_Color.pdf", plot = last_plot(), device = cairo_pdf, width = 3, height = 3)

```

#Determine the GC % for each barcode:
```{r, include=FALSE}

all_percentages_dataframe$GC_Content_Percent <- str_count(row.names(all_percentages_dataframe), "[GCgc]")*100/nchar(row.names(all_percentages_dataframe))

lm(formula = all_percentages_dataframe$Transduced ~ all_percentages_dataframe$gDNA + all_percentages_dataframe$GC_Content_Percent + all_percentages_dataframe$gDNA*all_percentages_dataframe$GC_Content_Percent)

```


```{r}
ggplot(all_percentages_dataframe,aes(x=GC_Content_Percent,y=Transduced))+geom_point()

ggplot(all_percentages_dataframe,aes(x=GC_Content_Percent,y=Transduced/gDNA))+geom_jitter()

```

#Plot box plot for % of cells/reads
```{r}

#Add lineage barcode as column for melting:
all_percentages_dataframe_w_bc <- all_percentages_dataframe[1:length(all_percentages_dataframe)-1]
all_percentages_dataframe_w_bc$Lineage_Barcode <- row.names(all_percentages_dataframe)

#Melt dataframe:
all_percentages_dataframe_melted <- melt(all_percentages_dataframe_w_bc)
colnames(all_percentages_dataframe_melted) <- c("Lineage_Barcode", "Source", "Percent")

ggplot(all_percentages_dataframe_melted, aes(x = Source, y = Percent)) +
  geom_boxplot(outlier.shape = NA) +
  geom_line(aes(group = Lineage_Barcode), alpha = 0.5, color = "gray60") +
  geom_point() +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    aspect.ratio = 1
  ) +
  xlab("") +
  ylab("% of Reads/Cells") +
  scale_x_discrete(limits = c("gDNA","Untransduced","Transduced"))


```
#Plot as gropuped bar plot plot, scaled:
```{r}


gDNA_transduced_only_percentages <- all_percentages_dataframe_melted[all_percentages_dataframe_melted$Source != "Untransduced",]

#Set factor to make untransduced visible along transduced:
scale_factor <- 60

ggplot(data = gDNA_transduced_only_percentages, 
       aes(x = Lineage_Barcode, y = ifelse(Source == "Transduced", Percent / scale_factor, Percent), fill = forcats::fct_rev(Source))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(
    name = "gDNA\n(% of Reads)",
    sec.axis = sec_axis(~ . * scale_factor, name = "Post-Enrichment Protocol\n(% of Cells)", breaks = seq(0,0.025*scale_factor, by = 0.005*scale_factor)),
    expand = expansion(mult = c(0,0.05))
  ) +
  scale_fill_manual(values = c("grey70", "black"),
                    labels = c("gDNA (% of Reads)",
                               "Post-Enrichment Protocol (% of Cells)")) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_blank(),
    legend.key.size = unit(0.4, "cm"),
    legend.spacing.y = unit(0.1, "cm"),
    legend.title = element_text(margin = margin(b=0.1, unit = "cm")),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  ) +
  xlab("Lineage Barcodes") +
  labs(fill = "Source") +
  coord_cartesian(ylim = c(0,0.025))

ggsave("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_Percentage_of_Cells/Barplot_Scaled.pdf", plot = last_plot(), device = cairo_pdf, width = 5.5, height = 1.5)

```
#Plot as gropuped bar plot plot, unscaled, focusing on gDNA, letting enriched go off scale:
```{r}


#Set factor to make untransduced visible along transduced:
scale_factor <- 1

ggplot(data = gDNA_transduced_only_percentages, 
       aes(x = Lineage_Barcode, y = ifelse(Source == "Transduced", Percent / scale_factor, Percent), fill = forcats::fct_rev(Source))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(
    name = "gDNA\n(% of Reads)",
    sec.axis = sec_axis(~ . * scale_factor, name = "Post-Enrichment Protocol\n(% of Cells)", breaks = seq(0,0.025*scale_factor, by = 0.005*scale_factor)),
    expand = expansion(mult = c(0,0.05))
  ) +
  scale_fill_manual(values = c("grey70", "black"),
                    labels = c("gDNA (% of Reads)",
                               "Post-Enrichment Protocol (% of Cells)")) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_blank(),
    legend.key.size = unit(0.4, "cm"),
    legend.spacing.y = unit(0.1, "cm"),
    legend.title = element_text(margin = margin(b=0.1, unit = "cm")),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  ) +
  xlab("Lineage Barcodes") +
  labs(fill = "Source") +
  coord_cartesian(ylim = c(0,0.025))

ggsave("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_Percentage_of_Cells/Barplot_Unscaled_gDNA.pdf", plot = last_plot(), device = cairo_pdf, width = 5.5, height = 1.5)


```


#Plot as gropuped bar plot plot, unscaled, focusing on enriched protocol:
```{r}


#Set factor to make untransduced visible along transduced:
scale_factor <- 1

ggplot(data = gDNA_transduced_only_percentages, 
       aes(x = Lineage_Barcode, y = ifelse(Source == "Transduced", Percent / scale_factor, Percent), fill = forcats::fct_rev(Source))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(
    name = "gDNA\n(% of Reads)",
    sec.axis = sec_axis(~ . * scale_factor, name = "Post-Enrichment Protocol\n(% of Cells)", breaks = seq(0,1.2*scale_factor, by = 0.4*scale_factor)),
    expand = expansion(mult = c(0,0.05))
  ) +
  scale_fill_manual(values = c("grey70", "black"),
                    labels = c("gDNA (% of Reads)",
                               "Post-Enrichment Protocol (% of Cells)")) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_blank(),
    legend.key.size = unit(0.4, "cm"),
    legend.spacing.y = unit(0.1, "cm"),
    legend.title = element_text(margin = margin(b=0.1, unit = "cm")),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  ) +
  xlab("Lineage Barcodes") +
  labs(fill = "Source") +
  coord_cartesian(ylim = c(0,1.5))

ggsave("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_Percentage_of_Cells/Barplot_Unscaled_Enriched.pdf", plot = last_plot(), device = cairo_pdf, width = 5.5, height = 1.5)


```
#Plot using a broken axis:

```{r}

ggplot(data = gDNA_transduced_only_percentages, 
       aes(x = Lineage_Barcode, y = ifelse(Source == "Transduced", Percent / scale_factor, Percent), fill = forcats::fct_rev(Source))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("grey70", "black"),
                    labels = c("gDNA (% of Reads)",
                               "Post-Enrichment Protocol (% of Cells)")) +
  xlab("Lineage Barcodes") +
  ylab("gDNA\n(% of Reads),\nEnrichment Library+\n(% of Cells)") +
  labs(fill = "Source") +
  # Add broken axis:
  scale_y_break(c(0.03, 0.04), scales = 1) +
  scale_y_break(c(0.2, 1.2), scales = 1) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_blank(),
    legend.key.size = unit(0.4, "cm"),
    legend.spacing.y = unit(0.1, "cm"),
    legend.title = element_text(margin = margin(b=0.1, unit = "cm")),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y.right = element_blank(),
    axis.title.y.right = element_blank()
  )

ggsave("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_Percentage_of_Cells/Barplot_Unscaled_Enriched_Broken_Axis.pdf", plot = last_plot(), device = cairo_pdf, width = 5.5, height = 2.2)

#Plot without legend for ease of transfer:

ggplot(data = gDNA_transduced_only_percentages, 
       aes(x = Lineage_Barcode, y = ifelse(Source == "Transduced", Percent / scale_factor, Percent), fill = forcats::fct_rev(Source))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("grey70", "black"),
                    labels = c("gDNA (% of Reads)",
                               "Post-Enrichment Protocol (% of Cells)")) +
  xlab("Lineage Barcodes") +
  ylab("gDNA\n(% of Reads),\nEnrichment Library+\n(% of Cells)") +
  labs(fill = "Source") +
  # Add broken axis:
  scale_y_break(c(0.03, 0.04), scales = 1) +
  scale_y_break(c(0.2, 1.2), scales = 1) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_blank(),
    legend.key.size = unit(0.4, "cm"),
    legend.spacing.y = unit(0.1, "cm"),
    legend.title = element_text(margin = margin(b=0.1, unit = "cm")),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y.right = element_blank(),
    axis.title.y.right = element_blank(),
    legend.position = "none"
  )


ggsave("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Plot_Percentage_of_Cells/Barplot_Unscaled_Enriched_Broken_Axis_NoLegend.pdf", plot = last_plot(), device = cairo_pdf, width = 3, height = 2.2)


```

#Add data to allow plotting non-zero cases in untransduced (since only 1 cell in 2 different lineages is present):
```{r, include=FALSE}

#Get barcodes that are nonzero in untransduced:
barcode_status <- all_percentages_dataframe_melted %>%
  filter(Source == "Untransduced") %>%
  mutate(Status = ifelse(Percent > 0, "nonzero", "zero")) %>%
  select(Lineage_Barcode, Status)

#Add untransduced to the dataframe:
untransduced_separate <- gDNA_transduced_only_percentages %>%
  left_join(barcode_status, by = "Lineage_Barcode") %>%
  mutate(Source = case_when(
    Source == "Transduced" & Status == "zero" ~ "Transduced_Untransduced_zero",
    Source == "Transduced" & Status == "nonzero" ~ "Transduced_Untransduced_nonzero",
    Source == "gDNA" & Status == "zero" ~ "gDNA_Untransduced_zero",
    Source == "gDNA" & Status == "nonzero" ~ "gDNA_Untransduced_nonzero",
    TRUE ~ Source
  )) %>%
  select(-Status)



```

#Create plot with untransduced:
```{r}


ggplot(data = untransduced_separate, 
       aes(x = Lineage_Barcode, y = Percent, fill = Source)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(breaks = c("gDNA_Untransduced_zero",
                               "Transduced_Untransduced_zero"),
                    values = c("gDNA_Untransduced_zero" = "grey70",
                               "Transduced_Untransduced_zero" = "black",
                               "gDNA_Untransduced_nonzero" = "pink",
                               "Transduced_Untransduced_nonzero" = "red"),
                    labels = c("gDNA (% of Reads)",
                               "Post-Enrichment Protocol (% of Cells)",
                               NULL,
                               NULL)) +
  xlab("Lineage Barcodes") +
  ylab("gDNA\n(% of Reads),\nEnrichment Library+\n(% of Cells)") +
  labs(fill = "Source") +
  # Add broken axis:
  scale_y_break(c(0.03, 0.04), scales = 1) +
  scale_y_break(c(0.2, 1.2), scales = 1) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_blank(),
    legend.key.size = unit(0.4, "cm"),
    legend.spacing.y = unit(0.1, "cm"),
    legend.title = element_text(margin = margin(b=0.1, unit = "cm")),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y.right = element_blank(),
    axis.title.y.right = element_blank()
  )

```
#Create plot with untransduced, plot with x labels to see order of barcodes:
```{r}


ggplot(data = untransduced_separate, 
       aes(x = Lineage_Barcode, y = Percent, fill = Source)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(breaks = c("gDNA_Untransduced_zero",
                               "Transduced_Untransduced_zero"),
                    values = c("gDNA_Untransduced_zero" = "grey70",
                               "Transduced_Untransduced_zero" = "black",
                               "gDNA_Untransduced_nonzero" = "pink",
                               "Transduced_Untransduced_nonzero" = "red"),
                    labels = c("gDNA (% of Reads)",
                               "Post-Enrichment Protocol (% of Cells)",
                               NULL,
                               NULL)) +
  xlab("Lineage Barcodes") +
  ylab("gDNA\n(% of Reads),\nEnrichment Library+\n(% of Cells)") +
  labs(fill = "Source") +
  # Add broken axis:
  scale_y_break(c(0.03, 0.04), scales = 1) +
  scale_y_break(c(0.2, 1.2), scales = 1) +
  theme_bw() +
  theme(
    text = element_text(family = font_type, size = font_size, color = "black"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white"),
    axis.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.title = element_text(family = font_type, size = font_size, color = "black"),
    legend.text = element_text(family = font_type, size = font_size, color = "black"),
    axis.text.x = element_text(angle = 90),
    legend.key.size = unit(0.4, "cm"),
    legend.spacing.y = unit(0.1, "cm"),
    legend.title = element_text(margin = margin(b=0.1, unit = "cm")),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y.right = element_blank(),
    axis.title.y.right = element_blank()
  )

```
