---
title: "Pseudobulk_Correlation"
output: html_document
date: "2026-02-09"
---

```{r setup, include=FALSE}

#Run only in chunk:
knitr::opts_knit$set(root.dir = '/Users/robertvandervelde/Library/CloudStorage/GoogleDrive-robertvandervelde@sydshafferlab.com/My Drive/Vander_Velde_Share/Experiments/RV143')

getwd()

```

#Load Libraries:
```{r, include=FALSE}

rm(list = ls())
library(Seurat)
library(tidyverse)
library(ggplot2)
library(tidyplots)
library(dplyr)
library(reshape2)

getwd()

set.seed(1)

```

#Set variable:
```{r, include=FALSE}

font_size <- 8
font_type <- "Helvetica Neue"

panel_dimensions <- c(20,20)

```

#Load seurat object:
```{r, include=FALSE}

seurat_strict_barcode_requirements <- readRDS("RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Organized_Data/Create_Seurat_Objects/seurat_strict_barcode_requirements")

```


#Get only cells in enrichment sample:
```{r, include=FALSE}

Idents(seurat_strict_barcode_requirements) <- "orig.ident"

seurat_strict_barcode_requirements_enrichment_only <- subset(x = seurat_strict_barcode_requirements, idents = "Enriched")
  
rm(seurat_strict_barcode_requirements)

```

#Isolate lineage with 3 or more cells:
```{r, include=FALSE}

#Get list of lineage with 3 or more cells:
lineage_list <- names(table(seurat_strict_barcode_requirements_enrichment_only$Mini_Library_Assignment)[table(seurat_strict_barcode_requirements_enrichment_only$Mini_Library_Assignment) >= 3])

#Remove "Non-Enriched" from lineage list:
lineage_list <- lineage_list[lineage_list != "Non-Enriched"]

Idents(seurat_strict_barcode_requirements_enrichment_only) <- "Mini_Library_Assignment"

#Subset based on list:
seurat_large_lineages <- subset(x = seurat_strict_barcode_requirements_enrichment_only, idents = lineage_list)

#Remove large seurat object:
rm(seurat_strict_barcode_requirements_enrichment_only)

```

#Create label list, with the number of cells in each lineage:
```{r, include=FALSE}

dictionary_labels <- table(seurat_large_lineages$Mini_Library_Assignment) %>% 
  {setNames(paste0("Lineage ", seq_along(.), " (cells = ", ., ")"), names(.))}

```

#Perform pseudobulk:
```{r, include=FALSE}

pseudobulk_object <- AggregateExpression(seurat_large_lineages)

```

#Threshold pseudobulk object:
```{r, include=FALSE}

pseudobulk_object$RNA_thresholded <- pseudobulk_object$RNA[rowSums(pseudobulk_object$RNA >= 5) == ncol(pseudobulk_object$RNA),]

```

#Normalize pseudobulked object:
```{r, include=FALSE}

pseudobulk_object$RNA_thresholded <- LogNormalize(pseudobulk_object$RNA_thresholded)

```
#Create correlation matrix:
```{r, include=FALSE}

#Use Spearman since linearity is not important:
correlation_matrix <- cor(as.data.frame(pseudobulk_object$RNA_thresholded), method = "spearman")

#Melt correlation matrix:
correlation_matrix_melted <- melt(correlation_matrix)

#Change order for plotting:
label_order <- names(dictionary_labels)[order(dictionary_labels)]

```

#Plot correlation plot:
```{r}

p <- tidyplot(data = correlation_matrix_melted, x = Var1, y = Var2, color = value) |>
  add_heatmap() |>
  adjust_size(width = panel_dimensions[1], height = panel_dimensions[2]) |>
  adjust_x_axis_title("") |>
  adjust_y_axis_title("") |>
  adjust_font(fontsize = font_size, family = font_type) |>
  adjust_x_axis(labels = dictionary_labels) |>
  adjust_y_axis(labels = dictionary_labels) |>
  reorder_x_axis_labels(label_order) |>
  reorder_y_axis_labels(rev(label_order))

ggsave(p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Pseudobulk_Correlation/Heatmap.pdf", device = cairo_pdf)

```

#Generate pseudoclones (random list of cells assigned to a clone, using the same number of cells as actual clones)
```{r, include=FALSE}

#Get cell numbers for random assignment:
table_cell_assignment <- table(seurat_large_lineages$Mini_Library_Assignment)

#Get list of all enriched cell barcodes:
temporary_list_cells <- colnames(seurat_large_lineages)

#Create dataframe for adding pseudolineages:
pseudolineage_df <- data.frame(Cell = temporary_list_cells, Pseudolineage = rep(NA, length(temporary_list_cells)))

#Loop across all lineage 'slots':

for (i in 1:length(table_cell_assignment)) {
  
  #Sample enough cells for current pseudolineage:
  sampled_cells <- sample(temporary_list_cells, as.numeric(table_cell_assignment[i]), replace = FALSE)
  
  #Add current pseudolineage score to those cells:
  pseudolineage_df$Pseudolineage[pseudolineage_df$Cell %in% sampled_cells] <- paste0("Pseudolineage ", i)
  
  #Remove sampled cells from temporary cell list:
  temporary_list_cells <- setdiff(temporary_list_cells, sampled_cells)
  
  #Remoove sampled cells variable for cleanliness:
  rm(sampled_cells)
  
}

row.names(pseudolineage_df) <- pseudolineage_df$Cell

pseudolineage_df <- pseudolineage_df[,2, drop = FALSE]

#Assign pseudolineage to Seurat object:
seurat_large_lineages <- AddMetaData(seurat_large_lineages,
            pseudolineage_df)

```

#Run pseudobulk analysis using pseudolineages:
```{r, include=FALSE}

Idents(seurat_large_lineages) <- "Pseudolineage"

pseudobulk_object_pseudolineages <- AggregateExpression(seurat_large_lineages)

#Threshold pseudolineage pseudobulk:
pseudobulk_object_pseudolineages$RNA_thresholded <- pseudobulk_object_pseudolineages$RNA[rowSums(pseudobulk_object_pseudolineages$RNA >= 5) == ncol(pseudobulk_object_pseudolineages$RNA),]

#Normalize pseudolineage pseudobulk:
pseudobulk_object_pseudolineages$RNA_thresholded <- LogNormalize(pseudobulk_object_pseudolineages$RNA_thresholded)

```

#Run correlation matrix on pseudobulked pseudolineages:
```{r, include=FALSE}

#Use Spearman since linearity is not important:
correlation_matrix_pseudolineages <- cor(as.data.frame(pseudobulk_object_pseudolineages$RNA_thresholded), method = "spearman")

#Melt correlation matrix:
correlation_matrix_pseudolineages_melted <- melt(correlation_matrix_pseudolineages)

#Double check number of cells in each pseudolineage:
table(seurat_large_lineages$Pseudolineage)x

#Create pseudolineage dictionary:

dictionary_labels_pseudolineage <- table(seurat_large_lineages$Pseudolineage) %>% 
  {setNames(paste0("Pseudolineage ", seq_along(.), " (cells = ", ., ")"), names(.))}

#Change order for plotting:
label_order_pseudolineage <- names(dictionary_labels_pseudolineage)[order(dictionary_labels_pseudolineage)]

```

#Plot correlation plot with pseudolineages:
```{r}

p <- tidyplot(data = correlation_matrix_pseudolineages, x = Var1, y = Var2, color = value) |>
  add_heatmap() |>
  adjust_size(width = panel_dimensions[1], height = panel_dimensions[2]) |>
  adjust_x_axis_title("") |>
  adjust_y_axis_title("") |>
  adjust_font(fontsize = font_size, family = font_type) |>
  adjust_x_axis(labels = dictionary_labels) |>
  adjust_y_axis(labels = dictionary_labels) |>
  reorder_x_axis_labels(label_order_pseudolineage) |>
  reorder_y_axis_labels(rev(label_order_pseudolineage))

ggsave(p, filename = "RV143_Comb_20250407_20250508_Seq/Post_Cluster_Analysis/Plots/Pseudobulk_Correlation/Heatmap_Pseudolineages.pdf", device = cairo_pdf)

```

